---
title: "16_gene_set_counts"
output: html_document
date: "2024-10-16"
---

# Counts per set 
This script adds up burden across a gene set.

For each person, the number of rare variants (allele count of 5 or less) is added up across all genes in the new NPD set from stage 15. This is done separately for high-confidence (LoFTEE) PTVs, deleterious (REVEL > 0.75) missense variants and synonymous variants. The input is the per-gene counts per person from stage 12 of the WES repository, which are split by chromosome (chromosome 1 is split into two halves). The output is one table per gene set and variant class, uploaded to the RAP. Counts for the other gene sets used (exome-wide and pLI ≥ 0.9) come from stage 14 of the WES repository. 

## Load packages
```{r}
library(data.table)
library(dplyr)
```
# Set variables 
```{r}
AC='5orless' # run this same script with this set differently to edit AC! 
chromosomes<-1:22
variant_classes<-list('HC_PTV', 'REVEL75_Miss', 'synonymous')

```
# Download data 
```{r}
# High-confidence PTVs
var<-'HC_PTV'
input_filename <- paste0('./Variant_counts/transposed_counts/AC', AC, '/AC', AC, '_', var,'/transposed*.tsv')
download_command <- paste('dx download ', shQuote(input_filename))
system(download_command)

# Pathogenic missense
var<-'REVEL75_Miss'
input_filename <- paste0('./Variant_counts/transposed_counts/AC', AC, '/AC', AC, '_', var,'/transposed*.tsv')
download_command <- paste('dx download ', shQuote(input_filename))
system(download_command)

# Synonymous variants
var<-'synonymous'
input_filename <- paste0('./Variant_counts/transposed_counts/AC', AC, '/AC', AC, '_', var,'/transposed*.tsv')
download_command <- paste('dx download ', shQuote(input_filename))
system(download_command)


# Gene set annotations
system('dx download ./ppt_bias_sets.csv')
sets=read.csv('ppt_bias_sets.csv', header=T)
```

# Neuropsychiatric disorder set
```{r}
setname='NPD_set_SCH_CCG_DD'
set=subset(sets, sets$NPD_set_SCH_CCG_DD==T) 
# Filter to only genes with a gene id (as this is what we'll match on)
list=set$gene_id


# Count
for (var in variant_classes){
  print(paste0('Running variant class ', var))
  for (chr in chromosomes) {
    # Process chr1 differently (as it's in 2 halves)
    if (chr==1) {
        # Read in 
        # Chr 1 halves are named h1/h2 for HC_PTV and first_half/second_half for other classes
        half_names <- if (var == 'HC_PTV') c('h1', 'h2') else c('first_half', 'second_half')
        first_half <- paste0('transposed_chr_1_', half_names[1], '_', var, '_AC', AC,'_gene_counts.tsv')
        first_half <- fread(first_half, header=TRUE, stringsAsFactors=FALSE)
        second_half <- paste0('transposed_chr_1_', half_names[2], '_', var, '_AC', AC,'_gene_counts.tsv')
        second_half <- fread(second_half, header=TRUE, stringsAsFactors=FALSE)

        # Merge to one table and add up the gene split across the two halves
        chr1 <- merge(first_half, second_half, by = "Sample_ID")
        chr1$ENSG00000198758 <- chr1$ENSG00000198758.x + chr1$ENSG00000198758.y
        chr1 <- chr1 %>% select(-ENSG00000198758.x, -ENSG00000198758.y)
        cat('n genes in chr 1:', length(chr1) - 1, '\n')
        
        # Filter to your gene set 
        chr1 <- chr1 %>% select(c("Sample_ID", any_of(list)))
        cat('n genes in chr 1 in set:', length(chr1) - 1, '\n')
        # Only proceed if there are genes in the filtered set
        if (ncol(chr1) > 1) {
            chr1[[paste0("chr_1_", var, "_AC", AC, "_total")]] <- rowSums(chr1[,-1])
            geneset <- select(chr1, c("Sample_ID", paste0("chr_1_", var, "_AC", AC, "_total")))
        } else {
            cat("No genes from the gene set found in chr 1.\n")
        }
        rm(chr1)
        rm(first_half)
        rm(second_half)
        gc()
    }
    else {
        # All other chromosomes
        filename <- paste0('transposed_chr_', chr, '_', var, '_AC', AC,'_gene_counts.tsv')
        input <- fread(filename, header=TRUE, stringsAsFactors=FALSE)
        print(paste0('Processing chr ', chr))
        cat('n genes in chr:', length(input) - 1, '\n')
        # Filter the columns to keep only the genes present in the gene list
        chr_set <- input %>% select(c("Sample_ID", any_of(list)))
        cat('n genes in set in chr:' , length(chr_set) -1, '\n')
        # Add up variants in genes in set 
        if (ncol(chr_set) > 1) {
            chr_set[[paste0("chr_", chr, "_", var, "_AC", AC, "_total")]] <- rowSums(chr_set[,-1])
            chr_set <- select(chr_set, c("Sample_ID", paste0("chr_", chr, "_", var, "_AC", AC, "_total")))
            
            # Merge to big table and print length (to show increasing in size!!)
            geneset <- merge(geneset, chr_set, by='Sample_ID', all=TRUE)  
            # 'all=TRUE' to retain all rows in case one file is missing data
            cat('n vars in set in chr:', sum(geneset[[paste0("chr_", chr, "_", var, "_AC", AC, "_total")]]), '\n')
        } else {
            cat("No genes from the gene set found in chr", chr, ".\n")
        }
        
        # Clear mem
        rm(input)
        rm(chr_set)
        gc()}
    }

  # Add up across all chromosomes 
  geneset[[paste0(var,"_AC", AC, "_", setname, "_total")]] <- rowSums(geneset[,-1])
  geneset <- select(geneset, c("Sample_ID", paste0(var,"_AC", AC, "_", setname, "_total")))
  print(sum(geneset[[2]]))

  # Save out 
  output <- paste0(setname, '_', var, '_AC', AC, '_counts.csv')
  write.csv(geneset, output)


  upload_command <- paste('dx upload ', output, '--destination ./Variant_counts/Gene_sets/')
  system(upload_command)
}
```

