---
title: "17_participation_phenotypes_processing"
output: html_document
date: "2024-09-24"
---

# Processing participation phenotypes
This script processes phenotype data from UKBB cognitive tests and participation measures. It was run in a mem1_ssd1_v2_x8 instance.

First, each person is marked as having completed each online cognitive test or not, in both rounds of online testing. Then z-scores are made for fluid intelligence and reaction time, which are used as measures of cognition. Next, the participation phenotypes are derived, such as whether someone completed any or all of the online cognitive tests, was lost to follow-up, or had an email address. Finally, flags for neuropsychiatric diagnoses (from ICD-10 codes) and the variables used to reweight the sample are added, and the script writes out `phenotypes_all_cleaned_inc_reweighting.csv` for stages 18 and 19.

# Set-up 
## Load required libraries 
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
install.packages('moments')
library(moments)
install.packages('corrplot')
library(corrplot)
library(purrr)

```

## Load in cognitive data
Note that all input files in this script come from the UK Biobank cohort browser. Required fields were selected there and exported to .csv using the Table Exporter tool on the RAP. 
```{r}
system('dx download ./phenotype_cohorts/*cog.csv')
online_cog=read.csv('online_cog.csv')
AC_cog=read.csv('AC_cog.csv')
cog=merge(AC_cog, online_cog, by='eid')

system('dx download ./phenotype_cohorts/missing_NM_phenotypes_participant.csv')
missing_phen=read.csv('missing_NM_phenotypes_participant.csv')
cog=merge(cog, missing_phen, by='eid')

system('dx download ./phenotype_cohorts/missing_phens_SDS_participant.csv')
missing_phen=read.csv('missing_phens_SDS_participant.csv')
cog=merge(cog, missing_phen, by='eid')

system('dx download ./phenotype_cohorts/prosp_mem.csv')
missing_phen=read.csv('prosp_mem.csv')
cog=merge(cog, missing_phen, by='eid')

cog$Sample_ID<-cog$eid
backup_cog<-cog

```

# Marking each online cognitive test as complete or not
## Numeric memory
This test was completed at in-person assessment centres and online. It is a test of working memory, and involved remembering numbers that progressively increased by one digit and inputting them immediately after presentation. 

### Online
```{r}
cog %>% count(p20240_i0)
cog$Online1_Numeric_memory_completed<- F
cog$Online1_Numeric_memory_completed[!is.na(cog$p20240_i0)]<- T
cog %>% count(Online1_Numeric_memory_completed)

```

## Pairs matching 
This test was completed at in-person assessment centres and online. It is a test of declarative memory. A matrix of face up cards was presented, and participants were asked to remember the position of as many pairs of matching cards as possible. Cards were then turned face-down, and participants selected the pairs of cards. Two rounds of this task were run, with the second involving a larger number of cards, and therefore pairs, to remember. 

### Online (only taken at online 1) 
```{r}
cog %>% count(p20244_i0)
cog$Online1_Pairs_matching_completed<- F
cog$Online1_Pairs_matching_completed[cog$p20244_i0=='Completed']<- T
cog %>% count(Online1_Pairs_matching_completed)
```
## TMTA
This test was completed at in-person assessment centres and online. It is a test of processing speed. Participants were presented with a set of digits scattered across the screen and asked to click on them in numeric order.
### Online - 1 
```{r}
cog %>% count(p20156_i0)
cog$Online1_TMTA_completed<- F
cog$Online1_TMTA_completed[!is.na(cog$p20156_i0)]<- T
cog %>% count(Online1_TMTA_completed)

```
### Online - 2 
```{r}
cog %>% count(p20156_i1)
cog$Online2_TMTA_completed<- F
cog$Online2_TMTA_completed[!is.na(cog$p20156_i1)]<- T
cog %>% count(Online2_TMTA_completed)
```

## TMTB
This test was completed at in-person assessment centres and online. It is a test of executive function. Participants were presented with a set of letters and digits scattered across the screen and were asked to click on them in alphanumeric order (e.g., A-1-B-2-C…).
### Online - 1 
```{r}
cog %>% count(p20157_i0)
cog$Online1_TMTB_completed<- F
cog$Online1_TMTB_completed[!is.na(cog$p20157_i0)]<- T
cog %>% count(Online1_TMTB_completed)

```
### Online - 2 
```{r}
cog %>% count(p20157_i1)
cog$Online2_TMTB_completed<- F
cog$Online2_TMTB_completed[!is.na(cog$p20157_i1)]<- T
cog %>% count(Online2_TMTB_completed)
```

## Fluid intelligence 
This test was completed at in-person assessment centres and online. It is a test of reasoning. Participants were presented with 13 questions designed to measure their problem solving, logic, and reasoning capacity, independent of acquired knowledge. They were given 2 minutes to answer as many multiple-choice questions as possible (with a maximum of 13). The questions got progressively harder. 
 
### Online - 1
```{r}
cog %>% count(p20191_i0)
cog$Online1_FI_completed<- F
cog$Online1_FI_completed[!is.na(cog$p20191_i0)]<- T
cog %>% count(Online1_FI_completed)
```

### Online - 2
```{r}
cog %>% count(p20191_i1)
cog$Online2_FI_completed<- F
cog$Online2_FI_completed[!is.na(cog$p20191_i1)]<- T
cog %>% count(Online2_FI_completed)
```


## Symbol digit substitution
This test was completed at in-person assessment centres and online. It is a test of processing speed. Participants were presented with a key of symbols matched to numbers and asked to indicate which number matched each symbol in a grid. 

### Online - 1
```{r}
cog %>% count(p20159_i0)
cog$Online1_SDS_completed<- F
cog$Online1_SDS_completed[!is.na(cog$p20159_i0)]<- T
cog %>% count(Online1_SDS_completed)
```

### Online - 2
```{r}
cog %>% count(p20159_i1)
cog$Online2_SDS_completed<- F
cog$Online2_SDS_completed[!is.na(cog$p20159_i1)]<- T
cog %>% count(Online2_SDS_completed)
```
## Matrix pattern completion
### Online - 2 (only taken here)
```{r}
cog %>% count(p20760_i1)
cog$Online2_MPC_completed<- F
cog$Online2_MPC_completed[!is.na(cog$p20760_i1)]<- T
cog %>% count(Online2_MPC_completed)
```

# Phenotypes reflecting cognitive ability
## RT 
Scores taken from baseline. This test was completed at in-person assessment centres and online. It is a test of processing speed. It is similar to a game of ‘snap’, played across 12 rounds. The participant was shown two cards and asked to press a button if they matched. Here, I select the mean time taken to press the button (UKBB field 20023) as the outcome measure. 
```{r}
cog %>% count(p20023_i0)

a=subset(cog, (!is.na(cog$p20023_i0)))
a=select(a, c("Sample_ID","p20023_i0"))
a$Reaction_time_AC1_mean_time_taken<-as.numeric(a$p20023_i0)

# Plot raw score
tiff("RT_AC1_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(Reaction_time_AC1_mean_time_taken))+
  geom_histogram(binwidth=20)+
  labs(x='Reaction time (first assessment centre) mean time to \n correctly identify matches (ms)')
dev.off()

# Transform 
a$Reaction_time_AC1_mean_time_taken_log = log(a$Reaction_time_AC1_mean_time_taken)

tiff("RT_AC1_log_transformed_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(Reaction_time_AC1_mean_time_taken_log))+
  geom_histogram(binwidth=0.05)+
  labs(x='Reaction time (first assessment centre) mean time to \n correctly identify matches (ms) - log transformed')
dev.off()

# Form z score
a$Reaction_time_AC1_mean_time_taken_Zscore=((a$Reaction_time_AC1_mean_time_taken_log
                                                    -mean(a$Reaction_time_AC1_mean_time_taken_log))
                                                   /sd(a$Reaction_time_AC1_mean_time_taken_log, na.rm = T))


# Remove outliers (+/- 4SDs of mean)
b=subset(a, a$Reaction_time_AC1_mean_time_taken_Zscore<4 & a$Reaction_time_AC1_mean_time_taken_Zscore>-4)
b$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm<-b$Reaction_time_AC1_mean_time_taken_Zscore
b<-select(b, "Sample_ID", "Reaction_time_AC1_mean_time_taken_Zscore_outliersrm")
a=merge(a, b, by='Sample_ID', all=T)


# Recode outliers(+/-4SDs of mean)
a$Reaction_time_AC1_mean_time_taken_Zscore_outliersrecoded<-a$Reaction_time_AC1_mean_time_taken_Zscore
a$Reaction_time_AC1_mean_time_taken_Zscore_outliersrecoded[a$Reaction_time_AC1_mean_time_taken_Zscore > 4] <- 4
a$Reaction_time_AC1_mean_time_taken_Zscore_outliersrecoded[a$Reaction_time_AC1_mean_time_taken_Zscore < -4] <- -4

a=select(a, c(1,3:7))
cog = merge(cog, a, by='Sample_ID', all = TRUE)

# Plot z score 
tiff("RT_AC1_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))+
  geom_histogram(binwidth=0.3)+
  labs(x='Reaction time (first assessment centre) mean time to \n correctly identify matches - Z score')
dev.off()
```
### Save plots out
```{r}
system('dx upload RT*.tiff --destination ./phenotype_cohorts/processed_cognition/')
```

## FI 
This test was completed at in-person assessment centres and online. It is a test of reasoning. Participants were presented with 13 questions designed to measure their problem solving, logic, and reasoning capacity, independent of acquired knowledge. They were given 2 minutes to answer as many multiple-choice questions as possible (with a maximum of 13). The questions got progressively harder. 
Here, the number of correct answers given (UKBB field 20016 and 20191) both in person and online as outcome measures.

## First assessment centre
```{r}
cog %>% count(p20016_i0)

a=subset(cog, !is.na(cog$p20016_i0))
a=select(a, c("Sample_ID","p20016_i0"))
a$FI_AC1_n_correct_answers<-as.numeric(a$p20016_i0)

# Plot raw score
tiff("FI_AC1_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(FI_AC1_n_correct_answers))+
  geom_histogram(binwidth=1)+
  labs(x='FI (first assessment centre) n correct answers')
dev.off()



# Form z score
a$FI_AC1_n_correct_answers_Zscore=((a$FI_AC1_n_correct_answers
                                                    -mean(a$FI_AC1_n_correct_answers))
                                                   /sd(a$FI_AC1_n_correct_answers, na.rm = T))
mean(a$FI_AC1_n_correct_answers_Zscore)
max(a$FI_AC1_n_correct_answers_Zscore)
min(a$FI_AC1_n_correct_answers_Zscore)

# No outlier removal needed

a=select(a, c(1,3:4))
cog = merge(cog, a, by='Sample_ID', all = TRUE)


# Plot z score 
tiff("FI_AC1_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(FI_AC1_n_correct_answers_Zscore))+
  geom_histogram(binwidth=0.5)+
  labs(x='FI (first assessment centre) n correct answers - Z score')
dev.off()
```



## Second assessment centre
```{r}
cog %>% count(p20016_i1)

a=subset(cog, !is.na(cog$p20016_i1))
a=select(a, c("Sample_ID","p20016_i1"))
a$FI_AC2_n_correct_answers<-as.numeric(a$p20016_i1)

# Plot raw score
tiff("FI_AC2_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(FI_AC2_n_correct_answers))+
  geom_histogram(binwidth=1)+
  labs(x='FI (second assessment centre) n correct answers')
dev.off()



# Form z score
a$FI_AC2_n_correct_answers_Zscore=((a$FI_AC2_n_correct_answers
                                                    -mean(a$FI_AC2_n_correct_answers))
                                                   /sd(a$FI_AC2_n_correct_answers, na.rm = T))
mean(a$FI_AC2_n_correct_answers_Zscore)
max(a$FI_AC2_n_correct_answers_Zscore)
min(a$FI_AC2_n_correct_answers_Zscore)

# No outlier removal needed


a=select(a, c(1,3:4))
cog = merge(cog, a, by='Sample_ID', all = TRUE)


# Plot z score 
tiff("FI_AC2_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(FI_AC2_n_correct_answers_Zscore))+
  geom_histogram(binwidth=0.9)+
  labs(x='FI (second assessment centre) n correct answers - Z score')
dev.off()
```

## Third assessment centre
```{r}
cog %>% count(p20016_i2)

a=subset(cog, !is.na(cog$p20016_i2))
a=select(a, c("Sample_ID","p20016_i2"))
a$FI_AC3_n_correct_answers<-as.numeric(a$p20016_i2)

# Plot raw score
tiff("FI_AC3_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(FI_AC3_n_correct_answers))+
  geom_histogram(binwidth=1)+
  labs(x='FI (third assessment centre) n correct answers')
dev.off()



# Form z score
a$FI_AC3_n_correct_answers_Zscore=((a$FI_AC3_n_correct_answers
                                                    -mean(a$FI_AC3_n_correct_answers))
                                                   /sd(a$FI_AC3_n_correct_answers, na.rm = T))
mean(a$FI_AC3_n_correct_answers_Zscore)
max(a$FI_AC3_n_correct_answers_Zscore)
min(a$FI_AC3_n_correct_answers_Zscore)

# No outlier removal needed


a=select(a, c(1,3:4))
cog = merge(cog, a, by='Sample_ID', all = TRUE)


# Plot z score 
tiff("FI_AC3_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(FI_AC3_n_correct_answers_Zscore))+
  geom_histogram(binwidth=0.9)+
  labs(x='FI (third assessment centre) n correct answers - Z score')
dev.off()
```

## Fourth assessment centre
```{r}
cog %>% count(p20016_i3)

a=subset(cog, !is.na(cog$p20016_i3))
a=select(a, c("Sample_ID","p20016_i3"))
a$FI_AC4_n_correct_answers<-as.numeric(a$p20016_i3)

# Plot raw score
tiff("FI_AC4_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(FI_AC4_n_correct_answers))+
  geom_histogram(binwidth=1)+
  labs(x='FI (fourth assessment centre) n correct answers')
dev.off()



# Form z score
a$FI_AC4_n_correct_answers_Zscore=((a$FI_AC4_n_correct_answers
                                                    -mean(a$FI_AC4_n_correct_answers))
                                                   /sd(a$FI_AC4_n_correct_answers, na.rm = T))
mean(a$FI_AC4_n_correct_answers_Zscore)
max(a$FI_AC4_n_correct_answers_Zscore)
min(a$FI_AC4_n_correct_answers_Zscore)

# No outlier removal needed


a=select(a, c(1,3:4))
cog = merge(cog, a, by='Sample_ID', all = TRUE)


# Plot z score 
tiff("FI_AC4_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(FI_AC4_n_correct_answers_Zscore))+
  geom_histogram(binwidth=0.5)+
  labs(x='FI (fourth assessment centre) n correct answers - Z score')
dev.off()

```

## Online - 1
```{r}
cog %>% count(p20191_i0)


a=subset(cog, !is.na(cog$p20191_i0))

a=select(a, c("Sample_ID","p20191_i0"))
# Select first time, so if someone completed in 2014 instance (i0) take this. If not, take 2017 instance (i1)
a$FI_online1_n_correct_answers <- a$p20191_i0
         




# Plot raw score
tiff("FI_online1_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(FI_online1_n_correct_answers))+
  geom_histogram(binwidth=1)+
  labs(x='FI (online 1) n correct answers')
dev.off()



# Form z score
a$FI_online1_n_correct_answers_Zscore=((a$FI_online1_n_correct_answers
                                                    -mean(a$FI_online1_n_correct_answers))
                                                   /sd(a$FI_online1_n_correct_answers, na.rm = T))
mean(a$FI_online1_n_correct_answers_Zscore)
max(a$FI_online1_n_correct_answers_Zscore)
min(a$FI_online1_n_correct_answers_Zscore)


a=select(a, c(1,3,4))
cog = merge(cog, a, by='Sample_ID', all = TRUE)


# Plot z score 
tiff("FI_online1_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(FI_online1_n_correct_answers_Zscore))+
  geom_histogram(binwidth=0.6)+
  labs(x='FI (online 1) n correct answers - Z score')
dev.off()

```


## Online - 2
```{r}
cog %>% count(p20191_i1)


a=subset(cog, !is.na(cog$p20191_i1))

a=select(a, c("Sample_ID","p20191_i1"))
a$FI_online2_n_correct_answers <- a$p20191_i1
         




# Plot raw score
tiff("FI_online2_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(a, aes(FI_online2_n_correct_answers))+
  geom_histogram(binwidth=1)+
  labs(x='FI (online 2) n correct answers')
dev.off()



# Form z score
a$FI_online2_n_correct_answers_Zscore=((a$FI_online2_n_correct_answers
                                                    -mean(a$FI_online2_n_correct_answers))
                                                   /sd(a$FI_online2_n_correct_answers, na.rm = T))
mean(a$FI_online2_n_correct_answers_Zscore)
max(a$FI_online2_n_correct_answers_Zscore)
min(a$FI_online2_n_correct_answers_Zscore)


a=select(a, c(1,3,4))
cog = merge(cog, a, by='Sample_ID', all = TRUE)




# Plot z score 
tiff("FI_online2_Zscore_distribution.tiff", width = 12, height = 9, units = "in", res = 300)
ggplot(cog, aes(FI_online2_n_correct_answers_Zscore))+
  geom_histogram(binwidth=0.6)+
  labs(x='FI (online 2) n correct answers - Z score')
dev.off()

```

## Save plots (and table) out
```{r}
system('dx upload FI*.tiff --destination ./phenotype_cohorts/processed_cognition/')
write.csv(cog, 'participation_cognition.csv')
system('dx upload participation_cognition.csv --destination ./phenotype_cohorts/processed_cognition/')
```

# Merge and derive more complex participation phenotypes 
```{r}
system('dx download ./phenotype_cohorts/participation_diagnoses_participant.csv')
phen=read.csv('participation_diagnoses_participant.csv')
system('dx download ./phenotype_cohorts/processed_cognition/participation_cognition.csv')
cog=read.csv('participation_cognition.csv')
phen=merge(phen, cog, by='eid')

```

# Participation phenotypes
## Death
```{r}
phen$date_of_death<-phen$p40000_i0
phen$date_of_death[phen$date_of_death==""]<-NA
phen$dead[is.na(phen$date_of_death)]<-F
phen$dead[!is.na(phen$date_of_death)]<-T
phen %>% count(dead)
```

## Email access 
```{r}
phen %>% count(p20005)
phen$email_access<-F
phen$email_access[phen$p20005=='Current, email address assumed working']<-T
phen %>% count(email_access) 
```

## Willing to attempt cognitive tests at baseline 
```{r}
phen %>% count(p62_i0)
phen$willing_to_attempt_cog_tests_at_baseline[phen$p62_i0=='Begin games']<-T
phen$willing_to_attempt_cog_tests_at_baseline[phen$p62_i0=='I am unable to try this']<-F

phen %>% count(willing_to_attempt_cog_tests_at_baseline) 
```


## Only interacted with UKBB at baseline
```{r}
# number of contacts 
phen$number_of_contacts_with_UKBB<-phen$p20145
phen %>% count(number_of_contacts_with_UKBB)

# Only contact at baseline 
phen$only_contact_with_UKBB_at_baseline_assessment[phen$number_of_contacts_with_UKBB==1]<-T
phen$only_contact_with_UKBB_at_baseline_assessment[phen$number_of_contacts_with_UKBB>1]<-F
phen %>% count(only_contact_with_UKBB_at_baseline_assessment) 

# Only contact at baseline considering death (removing all people who have died since baseline)
phen$only_contact_with_UKBB_at_baseline_assessment_dead_removed<-phen$only_contact_with_UKBB_at_baseline_assessment
phen$only_contact_with_UKBB_at_baseline_assessment_dead_removed[phen$dead==T]<-NA
phen%>% count(only_contact_with_UKBB_at_baseline_assessment_dead_removed)

# Only contact at baseline considering death and email access
phen$only_contact_with_UKBB_at_baseline_assessment_dead_removed_noemail_removed<-phen$only_contact_with_UKBB_at_baseline_assessment_dead_removed
phen$only_contact_with_UKBB_at_baseline_assessment_dead_removed_noemail_removed[phen$email_access==F]<-NA
```

## Online cognitive function tests
### Invitations 
First round (2014)
```{r}
phen$date_invited_to_online_cog_1 <- phen$p22017_i0
phen$date_invited_to_online_cog_1[phen$date_invited_to_online_cog_1==""]<-NA
phen$invited_to_online_cog_1[is.na(phen$date_invited_to_online_cog_1)]<-FALSE
phen$invited_to_online_cog_1[!is.na(phen$date_invited_to_online_cog_1)]<-TRUE 
phen %>% count(invited_to_online_cog_1)
```
Second round (2021/2022)
```{r}
phen$date_invited_to_online_cog_2 <- phen$p22017_i1
phen$date_invited_to_online_cog_2[phen$date_invited_to_online_cog_2==""]<-NA
phen$invited_to_online_cog_2[is.na(phen$date_invited_to_online_cog_2)]<-FALSE
phen$invited_to_online_cog_2[!is.na(phen$date_invited_to_online_cog_2)]<-TRUE 
phen %>% count(invited_to_online_cog_2) 
```
Both
```{r}
phen$invited_to_both_online_cog<- F
phen$invited_to_both_online_cog[phen$invited_to_online_cog_1==T &
                                  phen$invited_to_online_cog_2==T] <- T
phen %>% count(invited_to_both_online_cog)

```
#### Invitations and dates of death 
Tried to remove people whose 'date lost to follow-up' was before invitation to certain things, but some of them continued to participate after what UKBB classes as 'loss to follow-up', so now only considering death. 
```{r}
phen$date_of_death <- as.Date(phen$date_of_death)
a = subset(phen, !is.na(phen$date_of_death))
phen$date_invited_to_online_cog_1 <- as.Date(phen$date_invited_to_online_cog_1)
a = subset(phen, !is.na(phen$date_invited_to_online_cog_1))
phen$date_invited_to_online_cog_2 <- as.Date(phen$date_invited_to_online_cog_2)
a = subset(phen, !is.na(phen$date_invited_to_online_cog_2))

phen$death_before_online_cog_1 <- phen$date_of_death < phen$date_invited_to_online_cog_1
phen$death_before_online_cog_2 <- phen$date_of_death < phen$date_invited_to_online_cog_2

phen %>% count(death_before_online_cog_1)
phen %>% count(death_before_online_cog_2) 
```
Death before both
```{r}
phen$death_before_both_online_cog <- F
phen$death_before_both_online_cog[phen$death_before_online_cog_1== T |
                                    phen$death_before_online_cog_2 == T] <- T
phen %>% count(death_before_both_online_cog)
```
Invited and alive for both 
```{r}
phen$invited_and_alive_both_online_cog <- F
phen$invited_and_alive_both_online_cog[phen$invited_to_both_online_cog==T &
                                         phen$death_before_both_online_cog==F]<- T
phen %>% count(invited_and_alive_both_online_cog)
```
### Completed any online cog test
#### First round (2014)
This battery included FI, TMTA&B, SDS, PM, and NM 
```{r}
phen$completed_any_test_online_cog_1<-F
phen$completed_any_test_online_cog_1[(phen$Online1_FI_completed == T) | 
                                       (phen$Online1_Numeric_memory_completed == T) | 
                                       (phen$Online1_Pairs_matching_completed == T) |
                                       (phen$Online1_SDS_completed == T) | 
                                       (phen$Online1_TMTA_completed == T) |
                                       (phen$Online1_TMTB_completed == T)]<- T
phen %>% count(completed_any_test_online_cog_1)

# Considering invitation
phen$completed_any_test_online_cog_1_considering_invitation[phen$completed_any_test_online_cog_1==T]<-T
phen$completed_any_test_online_cog_1_considering_invitation[(phen$invited_to_online_cog_1 == T) &
                                                              (phen$completed_any_test_online_cog_1 == F)]<-F
phen %>% count(completed_any_test_online_cog_1_considering_invitation)

# Considering invitation and death 
phen$completed_any_test_online_cog_1_considering_invitation_and_death<-phen$completed_any_test_online_cog_1_considering_invitation
phen$completed_any_test_online_cog_1_considering_invitation_and_death[phen$death_before_online_cog_1 == T]<-NA
phen %>% count(completed_any_test_online_cog_1_considering_invitation_and_death)
phen %>% count(completed_any_test_online_cog_1_considering_invitation)
```

#### Second round (2022)
This battery included FI, TMTA&B, SDS, MPC
```{r}
phen$completed_any_test_online_cog_2<-F
phen$completed_any_test_online_cog_2[(phen$Online2_FI_completed == T) | 
                                       (phen$Online2_MPC_completed == T) | 
                                       (phen$Online2_SDS_completed == T) |
                                       (phen$Online2_TMTA_completed == T) | 
                                       (phen$Online2_TMTB_completed == T)]<- T
phen %>% count(completed_any_test_online_cog_2)

# Considering invitation
phen$completed_any_test_online_cog_2_considering_invitation[phen$completed_any_test_online_cog_2==T]<-T
phen$completed_any_test_online_cog_2_considering_invitation[phen$invited_to_online_cog_2==T&
                                                              phen$completed_any_test_online_cog_2==F]<-F
phen %>% count(completed_any_test_online_cog_2_considering_invitation)

# Considering invitiation and death 
phen$completed_any_test_online_cog_2_considering_invitation_and_death<-phen$completed_any_test_online_cog_2_considering_invitation
phen$completed_any_test_online_cog_2_considering_invitation_and_death[phen$death_before_online_cog_2==T]<-NA
phen %>% count(completed_any_test_online_cog_2_considering_invitation_and_death)
phen %>% count(completed_any_test_online_cog_2_considering_invitation)

```

#### Online 1 and 2 together 
```{r}
phen$completed_any_test_online_cog_either<-F
phen$completed_any_test_online_cog_either[phen$completed_any_test_online_cog_1==T |
                                            phen$completed_any_test_online_cog_2==T]<-T
phen %>% count(completed_any_test_online_cog_either)

# Considering invitation and death

phen$completed_any_test_online_cog_either_considering_invitation_and_death[
    phen$completed_any_test_online_cog_1_considering_invitation_and_death==F &
    phen$completed_any_test_online_cog_2_considering_invitation_and_death==F]<-F
phen$completed_any_test_online_cog_either_considering_invitation_and_death[
  phen$invited_and_alive_both_online_cog==F]<-NA
phen$completed_any_test_online_cog_either_considering_invitation_and_death[
    phen$completed_any_test_online_cog_1_considering_invitation_and_death==T |
    phen$completed_any_test_online_cog_2_considering_invitation_and_death==T]<-T

phen %>% count(completed_any_test_online_cog_either_considering_invitation_and_death)


```
### Completed all online cog test
#### First round (2014)
This battery included FI, TMTA&B, SDS, PM, and NM 
```{r}
phen$completed_all_test_online_cog_1<-F
phen$completed_all_test_online_cog_1[(phen$Online1_FI_completed == T) &
                                       (phen$Online1_Numeric_memory_completed == T) &
                                       (phen$Online1_Pairs_matching_completed == T) &
                                       (phen$Online1_SDS_completed == T) &
                                       (phen$Online1_TMTA_completed == T) &
                                       (phen$Online1_TMTB_completed == T)]<- T
phen %>% count(completed_all_test_online_cog_1)

# Considering invitation
phen$completed_all_test_online_cog_1_considering_invitation[phen$completed_all_test_online_cog_1==T]<-T
phen$completed_all_test_online_cog_1_considering_invitation[(phen$invited_to_online_cog_1 == T) &
                                                              (phen$completed_all_test_online_cog_1 == F)]<-F
phen %>% count(completed_all_test_online_cog_1_considering_invitation)

# Considering invitation and death 
phen$completed_all_test_online_cog_1_considering_invitation_and_death<-phen$completed_all_test_online_cog_1_considering_invitation
phen$completed_all_test_online_cog_1_considering_invitation_and_death[phen$death_before_online_cog_1 == T]<-NA
phen %>% count(completed_all_test_online_cog_1_considering_invitation_and_death)
```

#### Second round (2022)
This battery included FI, TMTA&B, SDS, MPC
```{r}
phen$completed_all_test_online_cog_2<-F
phen$completed_all_test_online_cog_2[(phen$Online2_FI_completed == T) &
                                       (phen$Online2_MPC_completed == T) &
                                       (phen$Online2_SDS_completed == T) &
                                       (phen$Online2_TMTA_completed == T) & 
                                       (phen$Online2_TMTB_completed == T)]<- T
phen %>% count(completed_all_test_online_cog_2)

# Considering invitation
phen$completed_all_test_online_cog_2_considering_invitation[phen$completed_all_test_online_cog_2==T]<-T
phen$completed_all_test_online_cog_2_considering_invitation[phen$invited_to_online_cog_2==T&
                                                              phen$completed_all_test_online_cog_2==F]<-F
phen %>% count(completed_all_test_online_cog_2_considering_invitation)

# Considering invitiation and death 
phen$completed_all_test_online_cog_2_considering_invitation_and_death<-phen$completed_all_test_online_cog_2_considering_invitation
phen$completed_all_test_online_cog_2_considering_invitation_and_death[phen$death_before_online_cog_2==T]<-NA
phen %>% count(completed_all_test_online_cog_2_considering_invitation_and_death)

```

#### Online 1 and 2 together 
```{r}
phen$completed_all_test_online_cog_both<-F
phen$completed_all_test_online_cog_both[phen$completed_all_test_online_cog_1==T &
                                            phen$completed_all_test_online_cog_2==T]<-T
phen %>% count(completed_all_test_online_cog_both)

# Considering invitation and death
phen$completed_all_test_online_cog_either_considering_invitation_and_death[
    phen$completed_all_test_online_cog_1_considering_invitation_and_death==F |
    phen$completed_all_test_online_cog_2_considering_invitation_and_death==F]<-F
phen$completed_all_test_online_cog_either_considering_invitation_and_death[
  phen$invited_and_alive_both_online_cog==F]<-NA
phen$completed_all_test_online_cog_either_considering_invitation_and_death[
    phen$completed_all_test_online_cog_1_considering_invitation_and_death==T &
    phen$completed_all_test_online_cog_2_considering_invitation_and_death==T]<-T
phen %>% count(completed_all_test_online_cog_either_considering_invitation_and_death)


```

## Save out
```{r}
write.csv(phen, 'phenotypes_cog_diagnoses_participation_7AUG25.csv')
system('dx upload phenotypes_cog_diagnoses_participation_7AUG25.csv')
```


# Diagnoses for sensitivity analyses
Broad 'neuropsychiatric disorder' phenotype, spanning ICD10 codes F20-29, F70-79, F84, and G40-41
## Read in data 
```{r}
system('dx download ./CACNA1C/data_participant.csv')
original_phenotypes=read.csv('data_participant.csv')
all_phen=merge(phen, original_phenotypes, by='eid')
```

F20s - SZ and related disorders
## F20 - SZ
```{r}
all_phen %>% count(p130874)
all_phen$F20_record <- F
all_phen$F20_record[all_phen$p130874!=""]<-T
all_phen %>% count(F20_record)
```

## F21 - schizoptypal
```{r}
all_phen %>% count(p130876)
all_phen$F21_record <- F
all_phen$F21_record[all_phen$p130876!=""]<-T
all_phen %>% count(F21_record)
```
## F22 - persistant delusional
```{r}
all_phen %>% count(p130878)
all_phen$F22_record <- F
all_phen$F22_record[all_phen$p130878!=""]<-T
all_phen %>% count(F22_record)
```

## F23 - psychotic disorders
```{r}
all_phen %>% count(p130880)
all_phen$F23_record <- F
all_phen$F23_record[all_phen$p130880!=""]<-T
all_phen %>% count(F23_record)
```

## F24 - induced delusional disorder
```{r}
all_phen %>% count(p130882)
all_phen$F24_record <- F
all_phen$F24_record[all_phen$p130882!=""]<-T
all_phen %>% count(F24_record)
```

## F25 - schizoaffective disorders
```{r}
all_phen %>% count(p130884)
all_phen$F25_record <- F
all_phen$F25_record[all_phen$p130884!=""]<-T
all_phen %>% count(F25_record)
```

## F28 - other non-organic psychosis
```{r}
all_phen %>% count(p130886)
all_phen$F28_record <- F
all_phen$F28_record[all_phen$p130886!=""]<-T
all_phen %>% count(F28_record)
```
## F29 - unspecified non-organic psychosis
```{r}
all_phen %>% count(p130888)
all_phen$F29_record <- F
all_phen$F29_record[all_phen$p130888!=""]<-T
all_phen %>% count(F29_record)
```

## F20s summary
```{r}
all_phen$F20s_summary = F
all_phen$F20s_summary[
    all_phen$F20_record==T |
    all_phen$F21_record==T |
    all_phen$F22_record==T | 
    all_phen$F23_record==T | 
    all_phen$F24_record==T |
    all_phen$F25_record==T |
    all_phen$F28_record==T | 
    all_phen$F29_record==T] <- T
all_phen %>% count(F20s_summary)
```
F70s - Intellectual disabilities
## F70 - mild mental retardation
```{r}
all_phen %>% count(p130950)
all_phen$F70_record <- F
all_phen$F70_record[all_phen$p130950!=""]<-T
all_phen %>% count(F70_record)
```

## F71 - moderate mental retardation
```{r}
all_phen %>% count(p130952)
all_phen$F71_record <- F
all_phen$F71_record[all_phen$p130952!=""]<-T
all_phen %>% count(F71_record)
```
## F72 - severe mental retardation 
```{r}
all_phen %>% count(p130954)
all_phen$F72_record <- F
all_phen$F72_record[all_phen$p130954!=""]<-T
all_phen %>% count(F72_record)
```

## F73 - profound mental retardation 


## F78 - other mental retardation
```{r}
all_phen %>% count(p130958)
all_phen$F78_record <- F
all_phen$F78_record[all_phen$p130958!=""]<-T
all_phen %>% count(F78_record)
```

## F79 - unspecified mental retardation
```{r}
all_phen %>% count(p130960)
all_phen$F79_record <- F
all_phen$F79_record[all_phen$p130960!=""]<-T
all_phen %>% count(F79_record)
```
## F70s summary
```{r}
all_phen$F70s_summary = F
all_phen$F70s_summary[
    all_phen$F70_record==T |
    all_phen$F71_record==T |
    all_phen$F72_record==T | 
    all_phen$F78_record==T | 
    all_phen$F79_record==T] <- T
all_phen %>% count(F70s_summary)
```

## F84 - pervasive developmental disorders 
```{r} 
all_phen %>% count(p130971)
all_phen$F84_record <-F
all_phen$F84_record[all_phen$p130971!='']<-T
all_phen %>% count(F84_record)
```
G40s - EPILEPSY AND RELATED DISORDERS
## G40 - EPILEPSY
```{r}
all_phen %>% count(p131048)
all_phen$G40_record <- F
all_phen$G40_record[all_phen$p131048!=""]<-T
all_phen %>% count(G40_record)
```

## G41 - STATUS EPILEPTICUS
```{r}
all_phen %>% count(p131050)
all_phen$G41_record <- F
all_phen$G41_record[all_phen$p131050!=""]<-T
all_phen %>% count(G41_record)
```

## Summary NPD field 
```{r}
all_phen$NPD_summary = F
all_phen$NPD_summary[
    all_phen$F20s_summary==TRUE |
    all_phen$F70s_summary==TRUE |
    all_phen$F84_record==TRUE |
    all_phen$G40_record==TRUE |
    all_phen$G41_record==TRUE] <- TRUE
all_phen %>% count(NPD_summary)

```

# Save out
```{r}
colnames(all_phen)
phen=all_phen[,c(1, 97:157, 239:257)]
write.csv(phen, 'phenotypes_all_cleaned_7AUG25.csv')
system('dx upload phenotypes_all_cleaned_7AUG25.csv')
```

# Reweighting 
```{r}
system('dx download ./phenotypes_all_cleaned_7AUG25.csv')
phen=read.csv('phenotypes_all_cleaned_7AUG25.csv')

system('dx download participation_bias/reweighting.csv')
weighting_phens=read.csv('reweighting.csv')
weighting_phens$housing_tenure = weighting_phens$p680_i0
weighting_phens$country_of_birth = weighting_phens$p1647_i0
weighting_phens$TDI = weighting_phens$p22189
weighting_phens$self_rep_ethnic_background = weighting_phens$p21000_i0
weighting_phens$breastfed_as_baby = weighting_phens$p1677_i0

weighting_phens=weighting_phens[,c(1, 10:14)]

phen=merge(phen, weighting_phens, by = 'eid')
write.csv(phen, 'phenotypes_all_cleaned_inc_reweighting.csv')
system('dx upload phenotypes_all_cleaned_inc_reweighting.csv --destination participation_bias/')
```

