---
title: "18_regressions"
output: html_document
date: "2025-08-26"
---
This script runs for <1 hour on a mem1_ssd1_v2_x16 RStudio instance.

This is the main analysis script. It reads in the phenotypes from stage 17, the gene set counts (from stage 16 here and stage 14 of the WES repository), pathogenic CNVs and covariates, and removes people with discordant sex. It then tests (using logistic regressions) whether carrying damaging rare coding variants (PTVs plus deleterious missense variants) or pathogenic CNVs is associated with each participation phenotype. Sensitivity analyses add neuropsychiatric diagnoses, cognitive ability, and online test invitation as covariates. Lastly, the script looks at how participation affects estimates of the effect of these variants on cognition (fluid intelligence and reaction time). This is done by running the regressions separately in people who did and didn't complete all online cognitive tests, first unweighted and then using inverse probability weights.

# Load libraries
```{r}
library(data.table)
library(dplyr)
library(purrr)
install.packages('broom')
library(broom)
library(ggplot2)
library(tidyr)
install.packages('car')
library(car)
install.packages("brglm2")
library(brglm2)
install.packages('rcompanion')
library(rcompanion)
install.packages('weights')
library(weights)
library(lubridate)
```

# Prep data
## Phenotypes
### Cognitive phenotypes
```{r}
system('dx download /participation_bias/phenotypes_all_cleaned_inc_reweighting.csv')
phen=read.csv('phenotypes_all_cleaned_inc_reweighting.csv', header=T)

# any instance
phen <- phen %>%
  mutate(
    FI_any = coalesce(
      FI_AC1_n_correct_answers,
      FI_AC2_n_correct_answers,
      FI_AC3_n_correct_answers,
      FI_AC4_n_correct_answers,
      FI_online1_n_correct_answers,
      FI_online2_n_correct_answers),
    FI_score_source = case_when(
      !is.na(FI_AC1_n_correct_answers) ~ "FI_AC1",
      !is.na(FI_AC2_n_correct_answers) ~ "FI_AC2",
      !is.na(FI_AC3_n_correct_answers) ~ "FI_AC3",
      !is.na(FI_AC4_n_correct_answers) ~ "FI_AC4",
      !is.na(FI_online1_n_correct_answers) ~ "FI_online1",
      !is.na(FI_online2_n_correct_answers) ~ "FI_online2",      
      TRUE ~ NA_character_ ))
phen$FI_any_zscore <- (phen$FI_any - mean(phen$FI_any, na.rm = TRUE)) / 
                           sd(phen$FI_any, na.rm = TRUE)

# Code RT so higher = better cognition
phen$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm <- (phen$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm)*(-1)

```

## Ancestries
Ancestry PCs are from stage 9 of the WES processing repository (https://github.com/eilidhfenner/UKBB_RAP_WES).
```{r}
system('dx download genetic_ancestries/ancestries_PCs.csv')
ancs=read.csv('ancestries_PCs.csv')
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')
rm(ancs)
gc()
```

## Genetic QC
Filter to those passing the final QC stage (sex imputation). The list of IDs with discordant sex is from stage 8 of the WES processing repository (https://github.com/eilidhfenner/UKBB_RAP_WES).
```{r}
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T
phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 
rm(ids_to_rm)
```

## Covariates
Read in covariates (and prep for use). These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)
covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared

system('dx download ./SZ_PRS/array_batch_participant.csv')
array_batch=read.csv('array_batch_participant.csv')
array_batch %>% count(array_batch$p22000)
array_batch=subset(array_batch, array_batch$p22000!='')
array_batch$genotype_batch<-array_batch$p22000
array_batch=array_batch[,c(1,3)]
covars=merge(covars, array_batch, by='eid')

phen=merge(phen, covars, by.x='eid', by.y='eid')
rm(covars)
rm(array_batch)
gc()
```

## Genetic data
### RCVs
Get counts
```{r}
AC='5orless'
system('dx download ./Variant_counts/Gene_sets/*_AC5orless_*.csv --overwrite')


file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)

# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames

# Merge
all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
```

### CNVs 
Pathogenic CNV calls are from Kendall et al. (2017, Biol. Psychiatry 82:103–110; 2019, Br. J. Psychiatry 214:297–304).
```{r}
system('dx download ./NDD_CNVs/Pathogenic_CNVs.csv')
CNVs=read.csv('Pathogenic_CNVs.csv')
CNVs$has_CNV_data<-T
CNVs %>% count(filter_good_CNVs)
CNVs$carries_HQ_CNV<-F
CNVs$carries_HQ_CNV[CNVs$filter_good_CNVs=='Selected']<-T

CNVs$SZ_enriched_CNV_carrier <- TRUE
CNVs$SZ_enriched_CNV_carrier[CNVs$carries_HQ_CNV==F]<-NA
CNVs %>% count(SZ_enriched_CNV_carrier)


CNVs_simple = CNVs[,c('eid','has_CNV_data', 
                      'carries_HQ_CNV', 'SZ_enriched_CNV_carrier')]

# Count n CNVs per person
cnv_counts <- CNVs_simple %>%
  group_by(eid) %>%
  summarise(n_CNVs = n())  # n() counts the number of rows per eid

# Diff field with just whether someone has any cnv or not 
CNVs_summary <- CNVs_simple %>%
  group_by(eid) %>%
  summarise(across(everything(), any)) 

# Add n cnvs to table 
CNVs_summary <- CNVs_summary %>%
  left_join(cnv_counts, by = "eid") 


combined = merge(all_ancs, CNVs_summary, by = 'eid', all.x=T) 

combined$n_CNVs_carried <- combined$n_CNVs
combined$n_CNVs_carried[is.na(combined$n_CNVs)] <- 0
combined$n_CNVs_carried[combined$carries_HQ_CNV==F]<-NA
combined %>% count(n_CNVs_carried)

all_ancs=subset(combined, !is.na(combined$n_CNVs_carried))


```


### RCV processing - add PTV+Miss RCVs 
```{r}
sets <- list('pLI_0.9',
             'NPD_set_SCH_CCG_DD')
library(stringr)

# Loop over each set name and create a new column for PTV + MIss
for (set_name in sets) {
  ptv_col <- paste0("HC_PTV_AC5orless_", set_name, "_total")
  miss_col <- paste0("REVEL75_Miss_AC5orless_", set_name, "_total")

  if (ptv_col %in% colnames(all_ancs) && miss_col %in% colnames(all_ancs)) {
    new_col <- paste0("HC_PTV_plus_REVEL75_Miss_AC5orless_", set_name, "_total")
    all_ancs[[new_col]] <- all_ancs[[ptv_col]] + all_ancs[[miss_col]]
  } else {
    warning(paste("Missing expected columns for set:", set_name))
    if (!(ptv_col %in% colnames(all_ancs))) message("  PTV column missing: ", ptv_col)
    if (!(miss_col %in% colnames(all_ancs))) message("  REVEL75_Miss column missing: ", miss_col)
  }
}
```

## Examine counts of variants and of phenotypes
```{r}
all_ancs %>% count(HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total)
all_ancs %>% count(HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total)
all_ancs %>% count(n_CNVs_carried)

all_ancs %>%
  count(willing_to_attempt_cog_tests_at_baseline) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

all_ancs %>%
  count(completed_any_test_online_cog_either) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

all_ancs %>%
  count(completed_all_test_online_cog_both) %>%
  mutate(percentage = round(100 * n / sum(n), 2))



# Cognitive tests
all_ancs %>% count(FI_any_zscore)
has_FI=subset(all_ancs, !is.na(all_ancs$FI_any_zscore))
has_FI %>% count(FI_score_source)
all_ancs %>% count(Reaction_time_AC1_mean_time_taken_Zscore_outliersrm)
has_RT=subset(all_ancs, !is.na(all_ancs$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
```
## Reweighting phenotypes
### Housing tenure
```{r}
all_ancs %>% count(housing_tenure)
all_ancs$housing_tenure[all_ancs$housing_tenure==""]<- NA

all_ancs <- all_ancs %>%
  mutate(
    housing_cat = case_when(
      housing_tenure %in% c("Own outright (by you or someone in your household)",
                            "Own with a mortgage",
                            "Pay part rent and part mortgage (shared ownership)") ~ "Own",
      housing_tenure %in% c("Rent - from local authority, local council, housing association",
                            "Rent - from private landlord or letting agency") ~ "Rent",
      housing_tenure %in% c("Live in accommodation rent free",
                            "None of the above") ~ "Shared/Other",
      housing_tenure == "Prefer not to answer" ~ "Prefer not to answer",
      TRUE ~ NA_character_
    ),
    housing_cat = factor(housing_cat, levels = c("Own","Rent","Shared/Other","Prefer not to answer"))
  )

all_ancs %>% count(housing_cat)

```

### Educational attainment
```{r}
system('dx download /participation_bias/additional_weighting_phenotypes.csv')
additional_weights=read.csv('additional_weighting_phenotypes.csv')
additional_weights$educational_attainment <- additional_weights$p6138_i0
additional_weights$educational_attainment[additional_weights$p6138_i0==""] <- NA
additional_weights %>% count(educational_attainment)
# Recode to highest educational attainment
levels_order <- c("Prefer not to answer", "None of the above",
                  "Other professional qualifications eg: nursing, teaching",
                  "NVQ or HND or HNC or equivalent",
                  "CSEs or equivalent",
                  "O levels/GCSEs or equivalent",
                  "A levels/AS levels or equivalent",
                  "College or University degree")
get_highest <- function(x) {
  quals <- unlist(strsplit(x, "\\|"))
  highest <- quals[which.max(match(quals, levels_order))]
  return(highest)
}
additional_weights$highest_education <- sapply(additional_weights$educational_attainment, get_highest)
additional_weights$highest_education <- factor(additional_weights$highest_education,
                                               levels = levels_order,
                                               ordered = TRUE)
all_ancs = merge(all_ancs, additional_weights, by = 'eid')
all_ancs %>% count(highest_education)

all_ancs <- all_ancs %>%
  mutate(
    education_cat = case_when(
      highest_education == "Prefer not to answer" ~ "Unknown",
      highest_education == "None of the above" ~ "Low",
      highest_education %in% c("Other professional qualifications eg: nursing, teaching","NVQ or HND or HNC or equivalent") ~ "Vocational/Other",
      highest_education %in% c("CSEs or equivalent","O levels/GCSEs or equivalent") ~ "Secondary education",
      highest_education == "A levels/AS levels or equivalent" ~ "A levels",
      highest_education == "College or University degree" ~ "Higher education"
    ),
    education_cat = factor(education_cat, 
                           levels = c("Unknown","Low","Vocational/Other","Secondary education","A levels","Higher education"),
                           ordered = TRUE)
  )

all_ancs %>% count(education_cat)
```

### Country of birth
```{r}
all_ancs %>% count(country_of_birth)
all_ancs$country_of_birth[all_ancs$country_of_birth==""]<- NA

all_ancs <- all_ancs %>%
  mutate(
    country_cat = case_when(
      country_of_birth == "England" ~ "England",
      country_of_birth %in% c("Scotland", "Wales", "Northern Ireland","Republic of Ireland") ~ "Other UK & ROI",
      country_of_birth == "Elsewhere" ~ "Elsewhere",
      country_of_birth %in% c("Do not know", "Prefer not to answer") ~ "Unknown",
      TRUE ~ NA_character_
    ),
    country_cat = factor(country_cat, levels = c("England","Other UK & ROI","Elsewhere","Unknown"))
  )

all_ancs %>% count(country_cat)

```

### TDI
```{r}
all_ancs %>% count(TDI)
```
### Self reported ethnic background
```{r}
all_ancs %>% count(self_rep_ethnic_background)
all_ancs$self_rep_ethnic_background[all_ancs$self_rep_ethnic_background==""]<- NA
all_ancs %>% count(self_rep_ethnic_background)

all_ancs <- all_ancs %>%
  mutate(
    ethnicity_cat = case_when(
      self_rep_ethnic_background %in% c("British","Irish","White","Any other white background") ~ "White",
      self_rep_ethnic_background %in% c("Indian","Pakistani","Bangladeshi","Any other Asian background", "White and Asian") ~ "South Asian or Asian-British",
      self_rep_ethnic_background %in% c("African","Caribbean","Any other Black background","Black or Black British", "White and Black African","White and Black Caribbean") ~ "Black or Black British",
      self_rep_ethnic_background %in% c("Mixed","Any other mixed background","Other ethnic group") ~ "Mixed/Other",
      self_rep_ethnic_background == "Chinese" ~ "Chinese",
      self_rep_ethnic_background %in% c("Do not know","Prefer not to answer") ~ "Unknown"
    ),
    ethnicity_cat = factor(ethnicity_cat, levels = c("White","South Asian or Asian-British","Black or Black British","Mixed/Other","Chinese","Unknown"))
  )

all_ancs %>% count(ethnicity_cat)
```

### Breastfed as a baby
```{r}
all_ancs %>% count(breastfed_as_baby)
all_ancs$breastfed_as_baby[all_ancs$breastfed_as_baby==""]<- NA
all_ancs %>% count(breastfed_as_baby)

all_ancs <- all_ancs %>%
  mutate(
    breastfed_cat = case_when(
      breastfed_as_baby == "Yes" ~ "Yes",
      breastfed_as_baby == "No" ~ "No",
      breastfed_as_baby %in% c("Do not know","Prefer not to answer") ~ "Unknown"
    ),
    breastfed_cat = factor(breastfed_cat, levels = c("Yes","No","Unknown"))
  )

all_ancs %>% count(breastfed_cat)
```

### Relationship status (taken from housing information)
```{r}
all_ancs$relationship <- grepl("Husband, wife or partner", all_ancs$p6141_i0)
all_ancs$relationship[(all_ancs$p709_i0=="")] <- NA
all_ancs %>% count(relationship)
all_ancs$relationship <- factor(all_ancs$relationship, levels = c(FALSE, TRUE))


```


### Age and sex
```{r}
all_ancs %>% count(standardised_age)
all_ancs %>% count(sex_1male)
all_ancs$sex_1male[is.na(all_ancs$sex_1male)] <- 0

```

### Employment
As people could answer >1, need a hierarchy: 
		Paid employment
		Student
		Home/family care
		Voluntary/Unpaid
		Sickness/disability
		Unemployed 
		Retired
		None of the above
		Prefer not to answer
```{r}
system('dx download participation_bias/employment_ACdate_participant.csv')
employment_acyear = read.csv('employment_ACdate_participant.csv')

all_ancs= merge(all_ancs, employment_acyear, by = 'eid')

all_ancs <- all_ancs %>%
  mutate(
    emp_raw = p6142_i0,
    employment_cat = case_when(
      grepl("paid employment|self-employed", p6142_i0, ignore.case = TRUE) ~ "Employed",
      grepl("student", p6142_i0, ignore.case = TRUE) ~ "Student",
      grepl("home and/or family", p6142_i0, ignore.case = TRUE) ~ "Home/family care",
      grepl("unpaid|voluntary", p6142_i0, ignore.case = TRUE) ~ "Voluntary/unpaid",
      grepl("sickness|disability", p6142_i0, ignore.case = TRUE) ~ "Disabled/Long-term sick",
      grepl("Unemployed", p6142_i0, ignore.case = TRUE) ~ "Unemployed",
      grepl("Retired", p6142_i0, ignore.case = TRUE) ~ "Retired",
      grepl("None", p6142_i0, ignore.case = TRUE) ~ "None of the above",
      grepl("Prefer", p6142_i0, ignore.case = TRUE) ~ "Prefer not to answer",
      TRUE ~ NA_character_
    ),
    employment_cat = factor(
      employment_cat,
      levels = c(
        "Employed",
        "Student",
        "Home/family care",
        "Voluntary/unpaid",
        "Disabled/Long-term sick",
        "Unemployed",
        "Retired",
        "None of the above",
        "Prefer not to answer"
      )
    )
  )

all_ancs %>% count(employment_cat)

all_ancs <- all_ancs %>%
  mutate(
    employment_collapsed = case_when(
      employment_cat == "Employed" ~ "Employed",
      employment_cat %in% c("Student","Home/family care","Voluntary/unpaid") ~ "Other",
      employment_cat == "Disabled/Long-term sick" ~ "Disabled/Long-term sick",
      employment_cat == "Unemployed" ~ "Unemployed",
      employment_cat == "Retired" ~ "Retired",
      employment_cat %in% c("None of the above","Prefer not to answer") ~ "Unknown"
    ),
    employment_collapsed = factor(
      employment_collapsed,
      levels = c("Employed","Other","Disabled/Long-term sick","Unemployed","Retired","Unknown")
    )
  )

all_ancs %>% count(employment_collapsed)
```

### Get subset of cohort who were alive at timepoint and with all reweighting phenotypes
```{r}
all_ancs$all_reweighting_phens <- F
all_ancs %>% count(death_before_both_online_cog)

all_ancs$all_reweighting_phens[
  !is.na(all_ancs$housing_cat) &
  !is.na(all_ancs$education_cat) &
  !is.na(all_ancs$country_cat) &
  !is.na(all_ancs$TDI) &
  !is.na(all_ancs$ethnicity_cat) &
  !is.na(all_ancs$relationship) &
  !is.na(all_ancs$breastfed_cat) &
  !is.na(all_ancs$standardised_age) &
  !is.na(all_ancs$sex_1male) &
  !is.na(all_ancs$employment_collapsed)&
  all_ancs$death_before_both_online_cog==F] <- T

all_ancs %>% count(all_reweighting_phens)

```


## Broad phens
```{r}
# Code broad participation phenotypes 
system('dx download participation_bias/broad_participation_participant.csv')
broad = read.csv('broad_participation_participant.csv')

# Lost to follow up 
broad %>% count(p20145)
broad$lost_to_follow_up <-0
broad$lost_to_follow_up[broad$p20145==1]<- 1
broad %>% count(lost_to_follow_up)

# No email access
broad %>% count(p20005)
broad$no_email_access <- 1
broad$no_email_access[broad$p20005=='Current, email address assumed working']<-0
broad %>% count(no_email_access)

# Did not complete MHQ
broad %>% count(p20400)
broad$no_MHQ <- 0
broad$no_MHQ[broad$p20400=='']<-1
broad %>% count(no_MHQ)


all_ancs = merge(all_ancs, broad, by='eid')


all_ancs %>%
  count(lost_to_follow_up) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

all_ancs %>%
  count(no_email_access) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

all_ancs %>%
  count(no_MHQ) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

```

## Clear mem
Can also remove files with RCV counts at this point. 
```{r}
rm(phen)
rm(covars)
rm(counts)
rm(additional_weights)
gc()
```

## Recode phenotypes 
1 = not participated, so that ORs make sense
```{r}
# Clean phens - recode binary phenotypes so they're all in the same direction: 1 = lower participation
dependent_var_pos<-c('completed_any_test_online_cog_either',
                     'completed_all_test_online_cog_both',
                     'willing_to_attempt_cog_tests_at_baseline')

dependent_var<-list('completed_any_test_online_cog_either',
                     'completed_all_test_online_cog_both',
                     'willing_to_attempt_cog_tests_at_baseline',
                     'no_MHQ',
                     'lost_to_follow_up',
                     'no_email_access')

# Convert these binary variables (currently listed as T/F into 0 and 1) - all now coded as 1 = did not participate and 0 = did participate 
all_ancs <- all_ancs %>%
  mutate(across(all_of(dependent_var_pos), ~ ifelse(. == TRUE, 0, ifelse(. == FALSE, 1, NA))))
all_ancs %>% count(completed_any_test_online_cog_either)

all_ancs %>% count(no_MHQ)
```

## Subset by genetically inferred ancestry 
```{r}
# Make subsets for diff ancs
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8)
EU_ancs_og <- EU_ancs
AAC_ancs=subset(all_ancs, all_ancs$AAC>0.8)
EAS_ancs=subset(all_ancs, all_ancs$EAS>0.8)
LAT_ancs=subset(all_ancs, all_ancs$LAT>0.8)
SAS_ancs=subset(all_ancs, all_ancs$SAS>0.8)
SSA_ancs=subset(all_ancs, all_ancs$SSA>0.8)
ADM_ancs=subset(all_ancs, (all_ancs$AAC<=0.8 & all_ancs$EAS<=0.8 & 
                              all_ancs$EUR<=0.8 & all_ancs$LAT<=0.8 & 
                              all_ancs$SAS<=0.8 & all_ancs$SSA<=0.8))

```

# BURDEN
## Regress - 1KGP-EUR-like ancestries
### Mean n vars
```{r}
mean(EU_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(EU_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(EU_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(EU_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(EU_ancs$n_CNVs_carried)
sd(EU_ancs$n_CNVs_carried)
```

### Prep results table 
```{r}
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_vars_0s = integer(),
  n_vars_1s = integer(),
  n_people = integer(),
  n_people_0s = integer(),
  n_people_1s = integer(),
  ancestry=character(),
  OR = numeric(),             
  lower_CI_OR = numeric(),    
  upper_CI_OR = numeric(),
  stringsAsFactors = FALSE
)
```

### RCVs
#### Set covars and var types
```{r}
# Check which ancestry probabilites have =< 5 non-0 values so can be useful covariates
EU_ancs %>% count(AAC) # Only 2 non-0 values, drop
EU_ancs %>% count(EAS) # Only 5 non-0 values, drop
EU_ancs %>% count(LAT) # Lots of non-0 values, keep as a useful covar
EU_ancs %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
EU_ancs %>% count(SSA) # 0 non-0 values, drop


covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```
Run

#### Regress 
##### Non Synonymous
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

```

##### Synonymous
```{r}

var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

```
run

### CNVs
#### Set covars and var types
```{r}
covars_Eu_anc_CNV <- 'AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch' 
```


#### Regress 
```{r}
var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
Run


#### Write out here in case of crashes
```{r}
write.csv(results, 'results_burden_EUR.csv')
system('dx upload results_burden_EUR.csv --destination participation_bias/inc_HCPTVs/')
```

## Mean n vars across ancestries
### AAC
```{r}
mean(AAC_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(AAC_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(AAC_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(AAC_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(AAC_ancs$n_CNVs_carried)
sd(AAC_ancs$n_CNVs_carried)
```

### Admixed
```{r}
mean(ADM_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(ADM_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(ADM_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(ADM_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(ADM_ancs$n_CNVs_carried)
sd(ADM_ancs$n_CNVs_carried)
```

### EAS
```{r}
mean(EAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(EAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(EAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(EAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(EAS_ancs$n_CNVs_carried)
sd(EAS_ancs$n_CNVs_carried)
```

### SAS
```{r}
mean(SAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(SAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(SAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(SAS_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(SAS_ancs$n_CNVs_carried)
sd(SAS_ancs$n_CNVs_carried)
```


### 1KGP-SSA-like ancestries
```{r}
mean(SSA_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(SSA_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(SSA_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(SSA_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(SSA_ancs$n_CNVs_carried)
sd(SSA_ancs$n_CNVs_carried)
```

### 1KGP-LAT-like ancestries
```{r}
mean(LAT_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
sd(LAT_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total, na.rm = T)
mean(LAT_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
sd(LAT_ancs$HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total, na.rm = T)
mean(LAT_ancs$n_CNVs_carried)
sd(LAT_ancs$n_CNVs_carried)
```

## Regress - 1KGP-SAS-like ancestries
Only non-EUR ancestry with the possibility of running without issue as largest. From here on, only the three participation in cognitive testing phenotypes are analysed.
```{r}
dependent_var<-list('completed_any_test_online_cog_either',
                    'completed_all_test_online_cog_both',
                    'willing_to_attempt_cog_tests_at_baseline')
```


### RCVs
#### Set covars and var types
```{r}
# Check which ancestry probabilites have =< 5 non-0 values so can be useful covariates
SAS_ancs %>% count(AAC) # Only 2 non-0 values, drop
SAS_ancs %>% count(EAS) # 0 non-0 values, drop
SAS_ancs %>% count(LAT) # 0 non-0 values, drop
SAS_ancs %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
SAS_ancs %>% count(SSA) # 0 non-0 values, drop


covars_SAS_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + SAS + SAS_PC1 + SAS_PC2 + SAS_PC3 + SAS_PC4 + SAS_PC5 + SAS_PC6 + SAS_PC7 + SAS_PC8 + SAS_PC9 + SAS_PC10' 

covars_SAS_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + SAS + SAS_PC1 + SAS_PC2 + SAS_PC3 + SAS_PC4 + SAS_PC5 + SAS_PC6 + SAS_PC7 + SAS_PC8 + SAS_PC9 + SAS_PC10'
```


#### Regress 
##### Non Synonymous
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SAS_anc_NS))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = SAS_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(SAS_ancs, !is.na(SAS_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "SAS_only",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

```
run

##### Synonymous
```{r}
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SAS_anc_syn))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = SAS_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(SAS_ancs, !is.na(SAS_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "SAS_only",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

```
run

### CNVs
```{r}
covars_SAS_anc_CNV <- 'AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + SAS + SAS_PC1 + SAS_PC2 + SAS_PC3 + SAS_PC4 + SAS_PC5 + SAS_PC6 + SAS_PC7 + SAS_PC8 + SAS_PC9 + SAS_PC10 + genotype_batch' 
var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_SAS_anc_CNV))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = SAS_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(SAS_ancs, !is.na(SAS_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "SAS_only",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
run

#### Write out here in case of crashes
```{r}
write.csv(results, 'results_burden_EUR_SAS.csv')
system('dx upload results_burden_EUR_SAS.csv --destination participation_bias/inc_HCPTVs/')
```


## Covarying for NPD diagnosis (1KGP-EUR-like ancestries)
### Count up
```{r}
all_ancs %>% count(NPD_summary)
EU_ancs %>% count(NPD_summary)
EU_ancs_NPD_removed = subset(EU_ancs, EU_ancs$NPD_summary==F)
```

### RCVs
#### Set covars and var types
```{r}
covars_Eu_anc_NS <- 'NPD_summary + seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' 

covars_Eu_anc_syn <- 'NPD_summary + seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```


#### Regress 
##### Non Synonymous
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_NPD_covar",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

```
run

##### Synonymous
```{r}
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_NPD_covar",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

# CNVs
covars_Eu_anc_CNV <- 'NPD_summary + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch' 

var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_NPD_covar",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
run

#### Write out here in case of crashes
```{r}
write.csv(results, 'results_burden_EUR_SAS_NPDcovar.csv')
system('dx upload results_burden_EUR_SAS_NPDcovar.csv --destination participation_bias/inc_HCPTVs/')
```

## Covary for cognition - 1KGP-EUR-like ancestries
### FI score 
#### Subset to just people with FI at baseline and then regress with and without FI covariate 
```{r}
EU_ancs_withFI = subset(EU_ancs, !is.na(EU_ancs$FI_AC1_n_correct_answers_Zscore))

# cant run with baseline cog testing phen, so drop this
dependent_var<-list('completed_any_test_online_cog_either',
                     'completed_all_test_online_cog_both')
```

#### Without covar
##### RCVs
###### Set covars and var types
```{r}
# Check which ancestry probabilites have =< 5 non-0 values so can be useful covariates
EU_ancs_withFI %>% count(AAC) # Only 2 non-0 values, drop
EU_ancs_withFI %>% count(EAS) # Only 5 non-0 values, drop
EU_ancs_withFI %>% count(LAT) # Lots of non-0 values, keep as a useful covar
EU_ancs_withFI %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
EU_ancs_withFI %>% count(SSA) # 0 non-0 values, drop


covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withFI, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withFI, !is.na(EU_ancs_withFI[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_FI_usual_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

# Synonymous
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withFI, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withFI, !is.na(EU_ancs_withFI[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_FI_usual_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

# CNVs
covars_Eu_anc_CNV <- 'AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch' 

var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs_withFI, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withFI, !is.na(EU_ancs_withFI[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_FI_usual_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
run


#### With covar for FI score
##### RCVs
###### Set covars and var types
```{r}
# Check which ancestry probabilites have =< 5 non-0 values so can be useful covariates
EU_ancs_withFI %>% count(AAC) # Only 2 non-0 values, drop
EU_ancs_withFI %>% count(EAS) # Only 5 non-0 values, drop
EU_ancs_withFI %>% count(LAT) # Lots of non-0 values, keep as a useful covar
EU_ancs_withFI %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
EU_ancs_withFI %>% count(SSA) # 0 non-0 values, drop


covars_Eu_anc_NS_FI <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + FI_AC1_n_correct_answers_Zscore' 

covars_Eu_anc_syn_FI <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + FI_AC1_n_correct_answers_Zscore'

# Non synonymous
var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS_FI))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withFI, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withFI, !is.na(EU_ancs_withFI[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_FI_FI_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}



# Synonymous

var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn_FI))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withFI, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withFI, !is.na(EU_ancs_withFI[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_FI_FI_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

# CNVs
covars_Eu_anc_CNV_FI <- 'AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch + FI_AC1_n_correct_answers_Zscore' 

var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV_FI))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs_withFI, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withFI, !is.na(EU_ancs_withFI[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_FI_FI_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
run

### RT score
#### Subset to just people with RT score and then regress with and without RT covariate for comparison
```{r}
EU_ancs_withRT = subset(EU_ancs, !is.na(EU_ancs$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
```

#### Without covar
##### RCVs
###### Set covars and var types
```{r}
# Check which ancestry probabilites have =< 5 non-0 values so can be useful covariates
EU_ancs_withRT %>% count(AAC) # Only 2 non-0 values, drop
EU_ancs_withRT %>% count(EAS) # Only 5 non-0 values, drop
EU_ancs_withRT %>% count(LAT) # Lots of non-0 values, keep as a useful covar
EU_ancs_withRT %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
EU_ancs_withRT %>% count(SSA) # 0 non-0 values, drop


covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' 

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'


# Regress - Non Synonymous

var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withRT, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withRT, !is.na(EU_ancs_withRT[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_RT_usual_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}


# Regress - Synonymous
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withRT, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withRT, !is.na(EU_ancs_withRT[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_RT_usual_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}


# CNVs
covars_Eu_anc_CNV <- 'AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch' 

var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs_withRT, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withRT, !is.na(EU_ancs_withRT[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_RT_usual_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
run

#### With covar for RT
##### RCVs
###### Set covars and var types
```{r}
# Check which ancestry probabilites have =< 5 non-0 values so can be useful covariates
EU_ancs_withRT %>% count(AAC) # Only 2 non-0 values, drop
EU_ancs_withRT %>% count(EAS) # Only 5 non-0 values, drop
EU_ancs_withRT %>% count(LAT) # Lots of non-0 values, keep as a useful covar
EU_ancs_withRT %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
EU_ancs_withRT %>% count(SSA) # 0 non-0 values, drop


covars_Eu_anc_NS_FI <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + Reaction_time_AC1_mean_time_taken_Zscore_outliersrm' 

covars_Eu_anc_syn_FI <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + Reaction_time_AC1_mean_time_taken_Zscore_outliersrm'


# Regress - Non Synonymous

var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS_FI))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withRT, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withRT, !is.na(EU_ancs_withRT[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_RT_RT_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}


# Regress - Synonymous
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn_FI))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs_withRT, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withRT, !is.na(EU_ancs_withRT[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_RT_RT_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

# CNVs
covars_Eu_anc_CNV_FI <- 'AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch + Reaction_time_AC1_mean_time_taken_Zscore_outliersrm' 

var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV_FI))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs_withRT, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs_withRT, !is.na(EU_ancs_withRT[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_with_RT_RT_covars",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

```
run

## Covarying for invitation
### Count up
```{r}
all_ancs %>% count(invited_and_alive_both_online_cog)
EU_ancs %>% count(invited_and_alive_both_online_cog)
```

### RCVs
#### Set covars and var types
```{r}
covars_Eu_anc_NS <- 'invited_and_alive_both_online_cog + seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10' 

covars_Eu_anc_syn <- 'invited_and_alive_both_online_cog + seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared  + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```


#### Regress 
##### Non Synonymous
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_invitation_covar",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}

```
run

##### Synonymous
```{r}
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var,'in ',set ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = 5,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_invitation_covar",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}}


# CNVs

covars_Eu_anc_CNV <- 'invited_and_alive_both_online_cog + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10 + genotype_batch' 
var_types<-list('n_CNVs_carried')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
      independent_var <- paste0(var)
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      print(paste('Running model testing association of ', dep_var, 'with ', independent_var ))
    
    model <- glm(formula, data = EU_ancs, family = binomial(),
        method = brglm2::brglmFit,
        type = "AS_mean")
    # Get coefficient summary
    coef_summary <- summary(model)$coefficients
    if (!(independent_var %in% rownames(coef_summary))) {
      message("Independent variable not found in model coefficients for ", dep_var, " and ", independent_var)
      next
    }
    beta     <- coef_summary[independent_var, "Estimate"]
    SE       <- coef_summary[independent_var, "Std. Error"]
    p_value  <- coef_summary[independent_var, "Pr(>|z|)"]
    conf_int <- confint(model)
    lower_CI <- conf_int[independent_var, 1]
    upper_CI <- conf_int[independent_var, 2]
    OR          <- exp(beta)
    lower_OR    <- exp(lower_CI)
    upper_OR    <- exp(upper_CI)
    # Get allele count and sample stats
    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]], na.rm = TRUE)
    with_phen_0s <- subset(with_phen, with_phen[[dep_var]] == FALSE)
    n_people_0s <- nrow(with_phen_0s)
    n_vars_0s <- sum(with_phen_0s[[independent_var]], na.rm = TRUE)
    with_phen_1s <- subset(with_phen, with_phen[[dep_var]] == TRUE)
    n_people_1s <- nrow(with_phen_1s)
    n_vars_1s <- sum(with_phen_1s[[independent_var]], na.rm = TRUE)
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = independent_var,
      allele_count = NA,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_vars_0s = n_vars_0s,
      n_vars_1s = n_vars_1s,
      n_people = n_people,
      n_people_0s = n_people_0s,
      n_people_1s = n_people_1s,
      ancestry = "EUR_only_invitation_covar",
      OR = OR,
      lower_CI_OR = lower_OR,
      upper_CI_OR = upper_OR,
      stringsAsFactors = FALSE
    ))
  }}

``` 
run


#### Write out
```{r}
write.csv(results, 'results_burden_all.csv')
system('dx upload results_burden_all.csv --destination participation_bias/inc_HCPTVs/')
```


# IMPACT
Subset cohort based on whether someone has completed ALL online cognitive tests and then run regression in each group separately. 

## Prep results table 
```{r}
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)
```

## Unweighted
### FI as outcome
#### Any online cognitive test 
##### Non-synonymous vars 
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')
dependent_var<-list('FI_any_zscore')
a=subset(EU_ancs_og, !is.na(EU_ancs_og$FI_any_zscore))

a %>% count(AAC) # Only 1 non-0 value, drop
a %>% count(EAS) # Only 4 non-0 values, drop
a %>% count(LAT) # Lots of non-0 values, keep as a useful covar
a %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
a %>% count(SSA) # 0 non-0 values, drop

covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

covars_Eu_anc_CNV <- 'genotype_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```
###### Completed ALL online cog tests
```{r}
a %>% count(completed_all_test_online_cog_both)
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1)

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```

##### Synonymous vars  
```{r}
var_types<-list('synonymous')

EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### CNVs
###### Completed ALL online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0) 

for (dep_var in dependent_var) {
  # Loop through each independent variable type
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }


```

###### Didn't
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }

```


### RT as outcome 
#### ALL online cognitive test 
##### Non-synonymous vars 
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')
dependent_var<-list('Reaction_time_AC1_mean_time_taken_Zscore_outliersrm')
a=subset(EU_ancs_og, !is.na(EU_ancs_og$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))

a %>% count(AAC) # Only 2 non-0 values, drop
a %>% count(EAS) # Only 5 non-0 values, drop
a %>% count(LAT) # Lots of non-0 values, keep as a useful covar
a %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
a %>% count(SSA) # 0 non-0 values, drop

# Therefore same covars as in FI 
```
###### Completed ALL online cog tests
```{r}
a %>% count(completed_all_test_online_cog_both)
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didn't
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1)

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```


##### Synonymous vars  
```{r}
var_types<-list('synonymous')

EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didn't
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### CNVs
###### Completed ALL online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0) 

for (dep_var in dependent_var) {
  # Loop through each independent variable type
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }

```


### Write out results 
```{r}
write.csv(results, 'results_FI_RT_regressions_AC5_unweighted_ALLonlinetests.csv')
system('dx upload results_FI_RT_regressions_AC5_unweighted_ALLonlinetests.csv --destination /participation_bias/inc_HCPTVs/')
```
## Weighted
### Prepare weights
Mirroring huang phenotypes plus participant vs parental
```{r}
all_ancs_weighting <- subset(all_ancs, all_ancs$all_reweighting_phens==T)

# Recode phenotypes for weighting
## Cog
# completed_all_test_online_cog_both was recoded above (1 = did not complete), so completers are 0 here. The weights model outcome is 1 = completed all online cognitive tests.
all_ancs_weighting$completed_all_online_cognitive_tests_weight_outcome <- 0
all_ancs_weighting$completed_all_online_cognitive_tests_weight_outcome[all_ancs_weighting$completed_all_test_online_cog_both == 0] <- 1
all_ancs_weighting %>% count(completed_all_online_cognitive_tests_weight_outcome)
```

### Derive weights 
```{r}
weights_model <- glm(completed_all_online_cognitive_tests_weight_outcome ~ housing_cat + 
                     education_cat + country_cat + TDI + ethnicity_cat + relationship +
                     breastfed_cat + standardised_age + sex_1male + employment_collapsed,
                   data = all_ancs_weighting, family = binomial())
a = tidy(weights_model)
summary(weights_model)

# Derive IP weights
all_ancs_weighting$prob_all_online <- predict(weights_model, type = "response")
all_ancs_weighting$nonresponse_weight_allonline <- 1 / all_ancs_weighting$prob_all_online
# Stabilise weights
all_ancs_weighting$sw_allonline <- mean(all_ancs_weighting$completed_all_online_cognitive_tests_weight_outcome) / all_ancs_weighting$prob_all_online

```

### Recode outliers
```{r}
outlier_cutoff <- mean(all_ancs_weighting$sw_allonline) + 4*(sd(all_ancs_weighting$sw_allonline))
all_ancs_weighting$sw_allonline_outliers <- ifelse(
  all_ancs_weighting$sw_allonline < outlier_cutoff,
  all_ancs_weighting$sw_allonline,
  outlier_cutoff)
range(all_ancs_weighting$sw_allonline_outliers)

``` 

### Look at the weights
```{r}
ggplot(all_ancs_weighting, aes(sw_allonline_outliers))+
  geom_histogram(bins = 100)+
  theme_minimal()

all_ancs_weighting$Group <- 
  factor(all_ancs_weighting$completed_all_online_cognitive_tests_weight_outcome,
         levels = c(0,1),
         labels = c("Did not complete all \n online cognitive tests", "Completed all online \n cognitive tests"))

ggplot(all_ancs_weighting, aes(sw_allonline_outliers, group = Group, colour = Group, fill = Group))+
  geom_histogram(alpha = 0.5, binwidth = 0.25)+
  theme_minimal()+
  labs(x = 'Stabilised IP weight',
       y = 'n')


EU_ancs_weighting = subset(all_ancs_weighting, (all_ancs_weighting$EUR > 0.8))
ggplot(EU_ancs_weighting, aes(sw_allonline_outliers, group = Group, colour = Group, fill = Group))+
  geom_histogram(alpha = 0.5, binwidth = 0.25)+
  theme_minimal()+
  labs(x = 'Stabilised IP weight',
       y = 'n')

``` 


### Weighted regressions - FI as outcome
##### Non-synonymous vars 
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')
dependent_var<-list('FI_any_zscore')
a=subset(EU_ancs_weighting, !is.na(EU_ancs_weighting$FI_any_zscore))

covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

covars_Eu_anc_CNV <- 'genotype_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```
###### Completed ALL online cog tests
```{r}
a %>% count(completed_all_test_online_cog_both)
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1)

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```


##### Synonymous vars  
```{r}
var_types<-list('synonymous')
```
###### Completed all online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### CNVs
###### Completed ALL online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0) 

for (dep_var in dependent_var) {
  # Loop through each independent variable type
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1)

# Loop through each phenotype
for (dep_var in dependent_var) {
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }

```


### Regressions weighting - RT as outcome 
#### Any online cognitive test 
##### Non-synonymous vars 
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')
dependent_var<-list('Reaction_time_AC1_mean_time_taken_Zscore_outliersrm')
a=subset(EU_ancs_weighting, !is.na(EU_ancs_weighting$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
```
###### Completed ALL online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1)

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

```


##### Synonymous vars  
```{r}
var_types<-list('synonymous')
```
###### Completed ALL online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0)

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE=SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = set,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}
```

##### CNVs
###### Completed ALL online cog tests
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==0) 

for (dep_var in dependent_var) {
  # Loop through each independent variable type
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_completed_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }


```

###### Didnt
```{r}
EU_ancs=subset(a, a$completed_all_test_online_cog_both==1) 

# Loop through each phenotype
for (dep_var in dependent_var) {
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "+", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs, weights = sw_allonline_outliers) 
      tidy_model <- tidy(model)

      # Extract outputs needed
      if (nrow(tidy_model) > 0) {
        beta <- tidy_model$estimate[2]
        SE <- tidy_model$std.error[2]
        conf_int <- confint(model)[2, ] 
        p_value <- tidy_model$p.value[2]
        with_phen=subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
        n_people<-nrow(with_phen)
        n_vars <- sum(with_phen[[independent_var]])
        anc_set<-'EU_anc_did_not_complete_all_online_tests_WEIGHTED'
        # Add to results data frame
        results <- rbind(results, data.frame(
          gene_set = NA,
          dep_var = dep_var,
          indep_var = independent_var,
          allele_count = AC,
          beta = beta,
          SE = SE,
          lower_CI = conf_int[1],
          upper_CI = conf_int[2],
          p_value = p_value,
          n_vars = n_vars,
          n_people = n_people,
          ancestry = anc_set,
          stringsAsFactors = FALSE
        ))
      }
    }

```

### Write out results 
```{r}
write.csv(results, 'results_FI_RT_regressions_AC5_incweighted.csv')
system('dx upload results_FI_RT_regressions_AC5_incweighted.csv --destination /participation_bias/inc_HCPTVs/')
```

