---
title: "19_interaction_analysis"
output: html_document
date: "2025-08-26"
---
This script runs for <1 hour on a mem1_ssd1_v2_x16 RStudio instance.

This script follows on from stage 18 and uses the same data preparation. Instead of splitting people into those who did and didn't complete all the online cognitive tests, an interaction is added between variant burden and completing the online tests. This tests directly whether the effect of damaging rare coding variants, synonymous variants or pathogenic CNVs on cognition (fluid intelligence and reaction time) differs by participation.

# Load libraries
```{r}
library(data.table)
library(dplyr)
library(purrr)
install.packages('broom')
library(broom)
library(ggplot2)
library(tidyr)
install.packages('car')
library(car)
install.packages("brglm2")
library(brglm2)
install.packages('rcompanion')
library(rcompanion)
install.packages('weights')
library(weights)
library(lubridate)
```


# Prep data
## Phenotypes
### Cognitive phenotypes
```{r}
system('dx download /participation_bias/phenotypes_all_cleaned_inc_reweighting.csv')
phen=read.csv('phenotypes_all_cleaned_inc_reweighting.csv', header=T)

# any instance
phen <- phen %>%
  mutate(
    FI_any = coalesce(
      FI_AC1_n_correct_answers,
      FI_AC2_n_correct_answers,
      FI_AC3_n_correct_answers,
      FI_AC4_n_correct_answers,
      FI_online1_n_correct_answers,
      FI_online2_n_correct_answers),
    FI_score_source = case_when(
      !is.na(FI_AC1_n_correct_answers) ~ "FI_AC1",
      !is.na(FI_AC2_n_correct_answers) ~ "FI_AC2",
      !is.na(FI_AC3_n_correct_answers) ~ "FI_AC3",
      !is.na(FI_AC4_n_correct_answers) ~ "FI_AC4",
      !is.na(FI_online1_n_correct_answers) ~ "FI_online1",
      !is.na(FI_online2_n_correct_answers) ~ "FI_online2",      
      TRUE ~ NA_character_ ))
phen$FI_any_zscore <- (phen$FI_any - mean(phen$FI_any, na.rm = TRUE)) / 
                           sd(phen$FI_any, na.rm = TRUE)

# Code RT so higher = better cognition
phen$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm <- (phen$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm)*(-1)

```

## Ancestries
Ancestry PCs are from stage 9 of the WES processing repository (https://github.com/eilidhfenner/UKBB_RAP_WES).
```{r}
system('dx download genetic_ancestries/ancestries_PCs.csv')
ancs=read.csv('ancestries_PCs.csv')
phen=merge(phen, ancs, by.x = 'eid', by.y='eid_13310')
rm(ancs)
gc()
```

## Genetic QC
Filter to those passing the final QC stage (sex imputation). The list of IDs with discordant sex is from stage 8 of the WES processing repository (https://github.com/eilidhfenner/UKBB_RAP_WES).
```{r}
system('dx download ./WES_QC/discordant_sex_ids.csv')
ids_to_rm=read.csv('discordant_sex_ids.csv')
ids_to_rm$remove=T
phen=merge(phen, ids_to_rm, by.x = 'eid', by.y='x', all=T)
phen=subset(phen, is.na(phen$remove)) 
rm(ids_to_rm)
```

## Covariates
Read in covariates (and prep for use). These phenotypes were selected in cohort browser and written out using the Table Exporter tool. 
```{r}
system('dx download ./phenotype_cohorts/covars_participant.csv')
covars=read.csv('covars_participant.csv', header=T)
covars$seq_batch<-covars$p32050
covars$AC_attended<-covars$p54_i0
covars$sex_1male[covars$p31 == 'Male']<-1
covars$sex_1male[covars$p31 == 'Female']<-2
covars$standardised_age <- (covars$p21022 - mean(covars$p21022, na.rm = TRUE)) / sd(covars$p21022, na.rm = TRUE)
covars$standardised_age_squared<-covars$standardised_age*covars$standardised_age
covars$sex_standardised_age<-covars$sex_1male * covars$standardised_age
covars$sex_standardised_age_squared<-covars$sex_1male * covars$standardised_age_squared

system('dx download ./SZ_PRS/array_batch_participant.csv')
array_batch=read.csv('array_batch_participant.csv')
array_batch %>% count(array_batch$p22000)
array_batch=subset(array_batch, array_batch$p22000!='')
array_batch$genotype_batch<-array_batch$p22000
array_batch=array_batch[,c(1,3)]
covars=merge(covars, array_batch, by='eid')

phen=merge(phen, covars, by.x='eid', by.y='eid')
rm(covars)
rm(array_batch)
gc()
```

## Genetic data
### RCVs
Get counts
```{r}
AC='5orless'
system('dx download ./Variant_counts/Gene_sets/*_AC5orless_*.csv --overwrite')

# REMOVE SINGLE GENE TABLES HERE!!
file_list <- list.files(path = "./", pattern = "*_AC5orless_counts.csv", full.names = TRUE)

# Read, drop first column (V1), and merge each file
counts <- lapply(file_list, function(file) {
  data <- fread(file)  # Read the file
  data[, V1 := NULL]  # Drop the first column (V1)
  return(data)
}) %>%
reduce(function(x, y) merge(x, y, by = "Sample_ID", all = TRUE))  # Merge the data frames

# Merge
all_ancs=merge(phen, counts, by.x='eid', by.y='Sample_ID') 
```

### CNVs 
Pathogenic CNV calls are from Kendall et al. (2017, Biol. Psychiatry 82:103–110; 2019, Br. J. Psychiatry 214:297–304).
```{r}
system('dx download ./NDD_CNVs/Pathogenic_CNVs.csv')
CNVs=read.csv('Pathogenic_CNVs.csv')
CNVs$has_CNV_data<-T
CNVs %>% count(filter_good_CNVs)
CNVs$carries_HQ_CNV<-F
CNVs$carries_HQ_CNV[CNVs$filter_good_CNVs=='Selected']<-T

CNVs$SZ_enriched_CNV_carrier <- TRUE
CNVs$SZ_enriched_CNV_carrier[CNVs$carries_HQ_CNV==F]<-NA
CNVs %>% count(SZ_enriched_CNV_carrier)


CNVs_simple = CNVs[,c('eid','has_CNV_data', 
                      'carries_HQ_CNV', 'SZ_enriched_CNV_carrier')]

# Count n CNVs per person
cnv_counts <- CNVs_simple %>%
  group_by(eid) %>%
  summarise(n_CNVs = n())  # n() counts the number of rows per eid

# Diff field with just whether someone has any cnv or not 
CNVs_summary <- CNVs_simple %>%
  group_by(eid) %>%
  summarise(across(everything(), any)) 

# Add n cnvs to table 
CNVs_summary <- CNVs_summary %>%
  left_join(cnv_counts, by = "eid") 


combined = merge(all_ancs, CNVs_summary, by = 'eid', all.x=T) 

combined$n_CNVs_carried <- combined$n_CNVs
combined$n_CNVs_carried[is.na(combined$n_CNVs)] <- 0
combined$n_CNVs_carried[combined$carries_HQ_CNV==F]<-NA
combined %>% count(n_CNVs_carried)

all_ancs=subset(combined, !is.na(combined$n_CNVs_carried))


```


### RCV processing - add PTV+Miss RCVs 
```{r}
sets <- list('pLI_0.9',
             'NPD_set_SCH_CCG_DD')
library(stringr)

# Loop over each set name and create a new column for PTV + MIss
for (set_name in sets) {
  ptv_col <- paste0("HC_PTV_AC5orless_", set_name, "_total")
  miss_col <- paste0("REVEL75_Miss_AC5orless_", set_name, "_total")

  if (ptv_col %in% colnames(all_ancs) && miss_col %in% colnames(all_ancs)) {
    new_col <- paste0("HC_PTV_plus_REVEL75_Miss_AC5orless_", set_name, "_total")
    all_ancs[[new_col]] <- all_ancs[[ptv_col]] + all_ancs[[miss_col]]
  } else {
    warning(paste("Missing expected columns for set:", set_name))
    if (!(ptv_col %in% colnames(all_ancs))) message("  PTV column missing: ", ptv_col)
    if (!(miss_col %in% colnames(all_ancs))) message("  REVEL75_Miss column missing: ", miss_col)
  }
}
```

## Examine counts of variants and of phenotypes
```{r}
all_ancs %>% count(HC_PTV_plus_REVEL75_Miss_AC5orless_pLI_0.9_total)
all_ancs %>% count(HC_PTV_plus_REVEL75_Miss_AC5orless_NPD_set_SCH_CCG_DD_total)
all_ancs %>% count(n_CNVs_carried)

all_ancs %>%
  count(willing_to_attempt_cog_tests_at_baseline) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

all_ancs %>%
  count(completed_any_test_online_cog_either) %>%
  mutate(percentage = round(100 * n / sum(n), 2))

all_ancs %>%
  count(completed_all_test_online_cog_both) %>%
  mutate(percentage = round(100 * n / sum(n), 2))



# Cognitive tests
all_ancs %>% count(FI_any_zscore)
has_FI=subset(all_ancs, !is.na(all_ancs$FI_any_zscore))
has_FI %>% count(FI_score_source)
all_ancs %>% count(Reaction_time_AC1_mean_time_taken_Zscore_outliersrm)
has_RT=subset(all_ancs, !is.na(all_ancs$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))
```

## Clear mem
Can also remove files with RCV counts at this point. 
```{r}
rm(phen)
rm(counts)
gc()
```

## Recode phenotypes 
1 = not participated, so that ORs make sense
```{r}
# Clean phens - recode binary phenotypes so they're all in the same direction: 1 = lower participation
dependent_var_pos<-c('completed_any_test_online_cog_either',
                     'completed_all_test_online_cog_both',
                     'willing_to_attempt_cog_tests_at_baseline')

dependent_var<-list('completed_any_test_online_cog_either',
                     'completed_all_test_online_cog_both',
                     'willing_to_attempt_cog_tests_at_baseline',
                     'no_MHQ',
                     'lost_to_follow_up',
                     'no_email_access')

# Convert these binary variables (currently listed as T/F into 0 and 1) - all now coded as 1 = did not participate and 0 = did participate 
all_ancs <- all_ancs %>%
  mutate(across(all_of(dependent_var_pos), ~ ifelse(. == TRUE, 0, ifelse(. == FALSE, 1, NA))))
all_ancs %>% count(completed_all_test_online_cog_both)

```

## Subset by genetically inferred ancestry 
```{r}
# Make subsets for diff ancs
EU_ancs=subset(all_ancs, all_ancs$EUR>0.8)
EU_ancs_og <- EU_ancs
AAC_ancs=subset(all_ancs, all_ancs$AAC>0.8)
EAS_ancs=subset(all_ancs, all_ancs$EAS>0.8)
LAT_ancs=subset(all_ancs, all_ancs$LAT>0.8)
SAS_ancs=subset(all_ancs, all_ancs$SAS>0.8)
SSA_ancs=subset(all_ancs, all_ancs$SSA>0.8)
ADM_ancs=subset(all_ancs, (all_ancs$AAC<=0.8 & all_ancs$EAS<=0.8 & 
                              all_ancs$EUR<=0.8 & all_ancs$LAT<=0.8 & 
                              all_ancs$SAS<=0.8 & all_ancs$SSA<=0.8))

```

## Set gene sets 
```{r}
sets <- list('pLI_0.9',
             'NPD_set_SCH_CCG_DD')


```

## Prep results table 
```{r}
results <- data.frame(
  gene_set=character(),
  dep_var = character(),
  indep_var = character(),
  allele_count = numeric(),
  beta = numeric(),
  SE = numeric(),
  lower_CI = numeric(),
  upper_CI = numeric(),
  p_value = numeric(),
  n_vars = integer(),
  n_people = integer(),
  ancestry=character(),
  stringsAsFactors = FALSE
)
```

## IMPACT
### FI as outcome
#### Non-synonymous vars 
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')
dependent_var<-list('FI_any_zscore')
a=subset(EU_ancs_og, !is.na(EU_ancs_og$FI_any_zscore))

a %>% count(AAC) # Only 1 non-0 value, drop
a %>% count(EAS) # Only 4 non-0 values, drop
a %>% count(LAT) # Lots of non-0 values, keep as a useful covar
a %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
a %>% count(SSA) # 0 non-0 values, drop

covars_Eu_anc_NS <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + synonymous_AC5orless_exome_wide_total + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

covars_Eu_anc_syn <- 'seq_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'

covars_Eu_anc_CNV <- 'genotype_batch + AC_attended + sex_1male + standardised_age + standardised_age_squared + sex_standardised_age + sex_standardised_age_squared + LAT + SAS + EU_PC1 + EU_PC2 + EU_PC3 + EU_PC4 + EU_PC5 + EU_PC6 + EU_PC7 + EU_PC8 + EU_PC9 + EU_PC10'
```

```{r}
EU_ancs=a

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "*completed_all_test_online_cog_both +", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

   terms_to_extract <- c(independent_var, 
                      "completed_all_test_online_cog_both", 
                      paste0(independent_var, ":completed_all_test_online_cog_both"))

    for (term in terms_to_extract) {
      if (term %in% tidy_model$term) {
    row <- tidy_model[tidy_model$term == term, ]
    conf <- confint(model)[term, ]
    
    beta <- row$estimate
    SE <- row$std.error
    p_value <- row$p.value
    lower_CI <- conf[1]
    upper_CI <- conf[2]

    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]])
    anc_set <- 'EU_anc'
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = paste0(independent_var, "_", term),  # label so you know which effect
      allele_count = AC,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_people = n_people,
      ancestry = anc_set,
      stringsAsFactors = FALSE
    ))
  }
}}}}


```


#### Synonymous vars  
```{r}
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "*completed_all_test_online_cog_both +", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

   terms_to_extract <- c(independent_var, 
                      "completed_all_test_online_cog_both", 
                      paste0(independent_var, ":completed_all_test_online_cog_both"))

    for (term in terms_to_extract) {
      if (term %in% tidy_model$term) {
    row <- tidy_model[tidy_model$term == term, ]
    conf <- confint(model)[term, ]
    
    beta <- row$estimate
    SE <- row$std.error
    p_value <- row$p.value
    lower_CI <- conf[1]
    upper_CI <- conf[2]

    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]])
    anc_set <- 'EU_anc'
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = paste0(independent_var, "_", term),  # label so you know which effect
      allele_count = AC,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_people = n_people,
      ancestry = anc_set,
      stringsAsFactors = FALSE
    ))
  }
}}}}
```

#### CNVs  
```{r}

for (dep_var in dependent_var) {
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "*completed_all_test_online_cog_both +", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

   terms_to_extract <- c(independent_var, 
                      "completed_all_test_online_cog_both", 
                      paste0(independent_var, ":completed_all_test_online_cog_both"))

    for (term in terms_to_extract) {
      if (term %in% tidy_model$term) {
    row <- tidy_model[tidy_model$term == term, ]
    conf <- confint(model)[term, ]
    
    beta <- row$estimate
    SE <- row$std.error
    p_value <- row$p.value
    lower_CI <- conf[1]
    upper_CI <- conf[2]

    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]])
    anc_set <- 'EU_anc'
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = paste0(independent_var, "_", term),  # label so you know which effect
      allele_count = AC,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_people = n_people,
      ancestry = anc_set,
      stringsAsFactors = FALSE
    ))
  }
    }}
```

### RT as outcome 
#### Non-synonymous vars 
```{r}
var_types<-list('HC_PTV_plus_REVEL75_Miss')
dependent_var<-list('Reaction_time_AC1_mean_time_taken_Zscore_outliersrm')
a=subset(EU_ancs_og, !is.na(EU_ancs_og$Reaction_time_AC1_mean_time_taken_Zscore_outliersrm))

a %>% count(AAC) # Only 2 non-0 values, drop
a %>% count(EAS) # Only 5 non-0 values, drop
a %>% count(LAT) # Lots of non-0 values, keep as a useful covar
a %>% count(SAS) #  Lots of non-0 values, keep as a useful covar
a %>% count(SSA) # 0 non-0 values, drop

# Therefore same covars as in FI 
```

```{r}
EU_ancs=a

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "*completed_all_test_online_cog_both +", covars_Eu_anc_NS))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

   terms_to_extract <- c(independent_var, 
                      "completed_all_test_online_cog_both", 
                      paste0(independent_var, ":completed_all_test_online_cog_both"))

    for (term in terms_to_extract) {
      if (term %in% tidy_model$term) {
    row <- tidy_model[tidy_model$term == term, ]
    conf <- confint(model)[term, ]
    
    beta <- row$estimate
    SE <- row$std.error
    p_value <- row$p.value
    lower_CI <- conf[1]
    upper_CI <- conf[2]

    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]])
    anc_set <- 'EU_anc'
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = paste0(independent_var, "_", term),  # label so you know which effect
      allele_count = AC,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_people = n_people,
      ancestry = anc_set,
      stringsAsFactors = FALSE
    ))
  }
}}}}


```


#### Synonymous vars  
```{r}
var_types<-list('synonymous')

for (dep_var in dependent_var) {
  # Loop through each independent variable type
  for (var in var_types) {
    # Loop through each set
    for (set in sets) {
      independent_var <- paste0(var, '_AC', AC, '_', set, '_total')
      formula <- as.formula(paste(dep_var, "~", independent_var, "*completed_all_test_online_cog_both +", covars_Eu_anc_syn))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

   terms_to_extract <- c(independent_var, 
                      "completed_all_test_online_cog_both", 
                      paste0(independent_var, ":completed_all_test_online_cog_both"))

    for (term in terms_to_extract) {
      if (term %in% tidy_model$term) {
    row <- tidy_model[tidy_model$term == term, ]
    conf <- confint(model)[term, ]
    
    beta <- row$estimate
    SE <- row$std.error
    p_value <- row$p.value
    lower_CI <- conf[1]
    upper_CI <- conf[2]

    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]])
    anc_set <- 'EU_anc'
    
    results <- rbind(results, data.frame(
      gene_set = set,
      dep_var = dep_var,
      indep_var = paste0(independent_var, "_", term),  # label so you know which effect
      allele_count = AC,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_people = n_people,
      ancestry = anc_set,
      stringsAsFactors = FALSE
    ))
  }
}}}}
```

#### CNVs  
```{r}

for (dep_var in dependent_var) {
      independent_var <- 'n_CNVs_carried'
      formula <- as.formula(paste(dep_var, "~", independent_var, "*completed_all_test_online_cog_both +", covars_Eu_anc_CNV))
      model <- lm(formula, data = EU_ancs) 
      tidy_model <- tidy(model)

   terms_to_extract <- c(independent_var, 
                      "completed_all_test_online_cog_both", 
                      paste0(independent_var, ":completed_all_test_online_cog_both"))

    for (term in terms_to_extract) {
      if (term %in% tidy_model$term) {
    row <- tidy_model[tidy_model$term == term, ]
    conf <- confint(model)[term, ]
    
    beta <- row$estimate
    SE <- row$std.error
    p_value <- row$p.value
    lower_CI <- conf[1]
    upper_CI <- conf[2]

    with_phen <- subset(EU_ancs, !is.na(EU_ancs[[dep_var]]))
    n_people <- nrow(with_phen)
    n_vars <- sum(with_phen[[independent_var]])
    anc_set <- 'EU_anc'
    
    results <- rbind(results, data.frame(
      gene_set = NA,
      dep_var = dep_var,
      indep_var = paste0(independent_var, "_", term),  # label so you know which effect
      allele_count = AC,
      beta = beta,
      SE = SE,
      lower_CI = lower_CI,
      upper_CI = upper_CI,
      p_value = p_value,
      n_vars = n_vars,
      n_people = n_people,
      ancestry = anc_set,
      stringsAsFactors = FALSE
    ))
  }
    }}
```

### Write out results 
```{r}
write.csv(results, 'results_FI_RT_interaction_regressions_AC5_unweighted.csv')
system('dx upload results_FI_RT_interaction_regressions_AC5_unweighted.csv --destination /participation_bias/inc_HCPTVs/')
```


